### 2021-05-18 writing/essay_latex/chapters/literature_review.tex
ported essay to latex

\chapter{Literature Review}
\label{chap:lit_review}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nullam feugiat lectus turpis. In quam arcu, faucibus quis nisl sodales, maximus venenatis ligula. Quisque suscipit nisi quis pretium interdum. Quisque placerat diam enim. Duis facilisis massa vitae diam volutpat dictum. Morbi diam dui, elementum eget tellus at, scelerisque egestas risus. Duis viverra sagittis congue. Curabitur laoreet consequat neque, eget venenatis odio commodo non. Cras sed volutpat velit, id ultrices quam. Pellentesque porttitor vehicula elit. Donec vitae lacus vel odio accumsan commodo sodales vel mi. Phasellus tortor leo, dignissim nec metus id, mattis aliquet ex.

Vestibulum ultrices dignissim lorem non finibus. Integer id mi vel ipsum condimentum sagittis varius at odio. Quisque laoreet nibh nec augue feugiat, eget vestibulum risus ultricies. Nunc ut sollicitudin leo. Ut fermentum risus quis varius mollis. Sed eu eros ut massa posuere porta eget in felis. Aliquam volutpat mollis consequat. Suspendisse potenti.

Integer aliquet elit quis ex aliquam, a sollicitudin neque dignissim. Suspendisse vestibulum quam mauris, sed consequat risus mollis vel. Nulla semper sem vitae lectus blandit vehicula. Nunc facilisis at lacus et malesuada. Phasellus id lacus vitae sapien pharetra vulputate. Proin bibendum quam ac diam hendrerit, id malesuada libero pharetra. Suspendisse finibus quam lobortis diam egestas, porttitor placerat leo bibendum. Sed aliquet dui vestibulum, rhoncus mi et, ornare arcu. In id nisl sed lorem fermentum euismod suscipit eget urna. Aenean sit amet lacinia lectus. Nunc iaculis libero ac tincidunt sodales.

Praesent scelerisque pretium libero, aliquam rhoncus orci auctor vel. Mauris eu lectus sed leo auctor imperdiet. Donec cursus nunc et est scelerisque, ac rutrum lorem accumsan. Morbi condimentum arcu et massa efficitur, et varius nisl rhoncus. Morbi rutrum ullamcorper ipsum. Aliquam vitae elementum nisl, sit amet molestie nunc. Curabitur tempor sit amet felis vitae egestas. Quisque quis risus ultrices, tincidunt tortor at, accumsan ligula. Curabitur vel nisi non nunc bibendum rhoncus eu nec arcu. Fusce et nisi id nulla tempor cursus vel eget ipsum.

Phasellus eu commodo metus, vitae pulvinar magna. Ut dignissim, nisl eu feugiat tempor, dui sem efficitur libero, sed dictum mauris justo ut urna. Donec ornare sapien ut lorem gravida, et accumsan nisi imperdiet. Mauris sem metus, volutpat a tempor feugiat, vestibulum ut urna. In mollis ultrices enim et tempor. Nullam enim libero, lobortis et pellentesque sodales, volutpat eu mi. Sed rhoncus vehicula rhoncus. Nullam in ultricies massa.

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nullam feugiat lectus turpis. In quam arcu, faucibus quis nisl sodales, maximus venenatis ligula. Quisque suscipit nisi quis pretium interdum. Quisque placerat diam enim. Duis facilisis massa vitae diam volutpat dictum. Morbi diam dui, elementum eget tellus at, scelerisque egestas risus. Duis viverra sagittis congue. Curabitur laoreet consequat neque, eget venenatis odio commodo non. Cras sed volutpat velit, id ultrices quam. Pellentesque porttitor vehicula elit. Donec vitae lacus vel odio accumsan commodo sodales vel mi. Phasellus tortor leo, dignissim nec metus id, mattis aliquet ex.

Vestibulum ultrices dignissim lorem non finibus. Integer id mi vel ipsum condimentum sagittis varius at odio. Quisque laoreet nibh nec augue feugiat, eget vestibulum risus ultricies. Nunc ut sollicitudin leo. Ut fermentum risus quis varius mollis. Sed eu eros ut massa posuere porta eget in felis. Aliquam volutpat mollis consequat. Suspendisse potenti.

Integer aliquet elit quis ex aliquam, a sollicitudin neque dignissim. Suspendisse vestibulum quam mauris, sed consequat risus mollis vel. Nulla semper sem vitae lectus blandit vehicula. Nunc facilisis at lacus et malesuada. Phasellus id lacus vitae sapien pharetra vulputate. Proin bibendum quam ac diam hendrerit, id malesuada libero pharetra. Suspendisse finibus quam lobortis diam egestas, porttitor placerat leo bibendum. Sed aliquet dui vestibulum, rhoncus mi et, ornare arcu. In id nisl sed lorem fermentum euismod suscipit eget urna. Aenean sit amet lacinia lectus. Nunc iaculis libero ac tincidunt sodales.

Praesent scelerisque pretium libero, aliquam rhoncus orci auctor vel. Mauris eu lectus sed leo auctor imperdiet. Donec cursus nunc et est scelerisque, ac rutrum lorem accumsan. Morbi condimentum arcu et massa efficitur, et varius nisl rhoncus. Morbi rutrum ullamcorper ipsum. Aliquam vitae elementum nisl, sit amet molestie nunc. Curabitur tempor sit amet felis vitae egestas. Quisque quis risus ultrices, tincidunt tortor at, accumsan ligula. Curabitur vel nisi non nunc bibendum rhoncus eu nec arcu. Fusce et nisi id nulla tempor cursus vel eget ipsum.

Phasellus eu commodo metus, vitae pulvinar magna. Ut dignissim, nisl eu feugiat tempor, dui sem efficitur libero, sed dictum mauris justo ut urna. Donec ornare sapien ut lorem gravida, et accumsan nisi imperdiet. Mauris sem metus, volutpat a tempor feugiat, vestibulum ut urna. In mollis ultrices enim et tempor. Nullam enim libero, lobortis et pellentesque sodales, volutpat eu mi. Sed rhoncus vehicula rhoncus. Nullam in ultricies massa.

### 2021-05-25 writing/essay_latex/draft3.tex
gh tweaks and tex progress

% sage_latex_guidelines.tex V1.20, 14 January 2017

\documentclass[sagev,times,Review]{sagej}
% add 'doublespace'

\usepackage{moreverb,url}

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\def\volumeyear{2021}

\begin{document}

\runninghead{Pape}

\title{Energy simulation and analysis of accessory dwelling unit: 
an evolutionary approach}

\author{Preston Pape\affilnum{1}}

\affiliation{\affilnum{1}University of Washington}

%\corrauth{}

\email{ppape@uw.edu}

\begin{abstract}
The United States is experiencing an unprecedented housing crisis, resulting in a rising population of unhoused peoples and the inability to become a homeowner or to even afford monthly rent in many cities. Solutions to this dilemma are neither straightforward nor definite. Seattle, Washington; Portland, Oregon; and Vancouver, British Columbia are exploring the use of accessory dwelling units (ADUs) as one method to combat surging housing prices. Use of ADUs is an effective means of increasing housing density without replacing single family housing zones with new multi family residential construction. Additionally, ADUs are often designed to be rented, generating supplementary income for the homeowner. Ten detached ADU (or DADU) designs are pre-approved by the City of Seattle and are freely available online to entice homeowners. However, a 2019 city survey shows that there are calls for an increased focus on sustainability and cost. This research intends to explore whether the use of genetic algorithms via shape grammar methodology to optimize DADU plans to site context, increases building performance or further encourages construction. The proposed methodology begins by reading example site data from the city of Seattle including building and vegetation context and rental/land prices from Seattle GIS and Zillow, respectively. Next, a genetic algorithm explores the design space for a viable floor plan solution based on a fitness function. This fitness function evaluates individual designs according to predefined traits. Traits to evaluate include window to wall ratio, insulation depth/type, ventilation strategy, and shading technique. Locating the correct combination of traits to minimize (or maximize) which results in a higher performance DADU is the desired outcome. Resulting designs will be analyzed and compared via energy performance simulation. The end-goal is to develop a computational tool using the aforementioned system to conduct automated site analysis, as well as parametric generation of DADUs.\end{abstract}

\keywords{Genetic Algorithm, ADU, Optimization, Simulation, EUI, Grasshopper, Galapagos}

\maketitle

\section{Introduction}
The United States is amid an unprecedented housing crisis, stemming from issues outside of housing, such as neoliberal cuts to social benefits spending and the increasing privatization of essentially every aspect of American life. The result is a rising population of unhoused peoples and the inability for many to become a homeowner or to even afford minimum monthly rent in many cities. Solutions to this dilemma are neither straightforward nor definite. However, as designers and architects, this problem can only be addressed at the symptom-level, while advocating at the root cause.

%TODO Expand background portion of intro

Seattle, Washington; Portland, Oregon; and Vancouver, British Columbia are exploring the use of accessory dwelling units (ADUs) as one method to combat surging housing prices. Use of ADUs is an effective means of increasing housing density without replacing single family housing zones with new multi family residential construction. Additionally, ADUs are often designed to be rented, generating supplementary income for the homeowner. Ten detached ADU (or DADU) designs are pre-approved by the City of Seattle and are available online to entice homeowners. However, a 2019 city survey shows that there are calls for an increased focus on sustainability and cost in the construction of ADUs\cite{seattlePreapprovedPlansAccessory2019}. These out of the box designs do not offer the scalability or energy efficiency that an ADU designed specifically per site offers.

%TODO Expand reason for why this tool is necessary to drive ADU demand

Homeowner’s associations and other local organizations have historically fought back against any proposed density increases through zoning or other method. However, Seattle and the other aforementioned cities in the Pacific Northwest have succeeded in allowing for the construction of ADUs in recent decades. The pre-approved designs are free, but require payment of around one thousand dollars [check this] for approval and come with one significant downside of many other pre-designed structures- a lack of contextual design and individualization.

This research explores whether the use of genetic algorithms via shape grammar methodology can effectively optimize DADU plans to site context, increases building performance or further encourages construction. The proposed methodology begins by reading example site data from the city of Seattle including building and vegetation context from the Seattle GIS. Next, a genetic algorithm explores the design space for a viable floor plan solution based on a fitness function. This fitness function evaluates individual designs according to predefined traits. Traits to evaluate include window to wall ratio, insulation depth/type, ventilation strategy, and shading technique. Locating the correct combination of traits to minimize (or maximize) which results in a higher performance DADU is the desired outcome. Resulting designs will be analyzed and compared via energy performance simulation. The end-goal is to develop a computational tool using the aforementioned system to conduct automated site analysis, as well as parametric generation of DADUs.

\section{Literature Review}
%TODO include feedback info on information from seattle city ADU survey- see obsidian notes: city survey respondents said “»  **Successful designs respond to site and context.** Some respondents lauded design choices that reflect a site’s opportunities and constraints or the architecture and climate of the Pacific Northwest. “also: “**Various types of spaces are valuable.** Respondents expressed interest in designs that include patios, rooftop decks, storage, and garages. “ - **SPACES TO CONSIDER FOR ALGORITHM** also: **DIY opportunities.** Respondents encouraged designs that allow homeowners to do a substantial amount of the construction themselves. “ - **SELF BUILD//NO ARCHITECT/BUILDER INVOLVEMENT don’t say no to this yet!!! (family/friend labor?)**also: **Flexibility.** Respondents emphasized the desirability of allowing mix-and-match options, such as multiple roof options, that make the design somewhat customizable. “ - **SHAPE GRAMMAR?**

\section{Methods} 
\subsection{Design constraints and decisions}
%TODO code restrictions: height limitations, lot size, etc via aduniverse-seattlecitygis.hub.arcgis.com/pages/code
%TODO touch on how 3 varying sites were chosen in 3 neighborhoods: central district, first hill and madrona
%TODO include table of R values for each layer, as well as U value of wall assembly, description of wood frame construction chosen, and why chosen.
\subsection{Energy simulation and analysis}
\subsection{Genetic algorithm}

\section{Results}

\section{Conclusion}

\subsection{Future work}
Benefit of a pure python tool 
Looking forward, moving on from the Grasshopper platform gives the benefit of non-reliance on developers to maintain the software in which the tool depends. Creation of such a tool in a singular programming language (in this case Python) further offers the ability to quickly and easily run the tool on a high variance of devices. In turn, this theoretically increases the rate of adoption by lowering the requirement to use the design tool. Additionally, Python is used within Rhino/Grasshopper (or a flavor thereof), provides many useful math and geometry libraries, and has options for injection into a web app (Flask and Django).
Limitations of Galapagos/Grasshopper workflow
Utilizing Galapagos and Grasshopper offers a ‘sandbox’ environment to begin to understand genetic algorithms and to arrive at tangible design solutions faster than using a home-brewed algorithm. Galapagos offers many fewer input parameters and a much narrower scope in which to define the fitness function. In its out-of-the-box form, Galapagos only accepts number sliders as input and can only optimize integers and floats- in reality, there is not means in which to define a true fitness function, only numerical values to target.

Figures are called in as follows:
\begin{verbatim}
\begin{figure}
\centering
\includegraphics{<figure name>}
\caption{<Figure caption>}
\end{figure}
\end{verbatim}

%\begin{figure}
%\setlength{\fboxsep}{0pt}%
%\setlength{\fboxrule}{0pt}%
%\begin{center}
%\begin{boxedverbatim}
%\begin{table}
%\small\sf\centering
%\caption{<Table caption.>}
%\begin{tabular}{<table alignment>}
%\toprule
%<column headings>\\
%\midrule
%<table entries
%(separated by & as usual)>\\
%<table entries>\\
%.
%.
%.\\
%\bottomrule
%\end{tabular}
%\end{table}
%\end{boxedverbatim}
%\end{center}
%\caption{Example table layout.\label{F2}}
%\end{figure}

%\subsection{Cross-referencing}
%The use of the \LaTeX\ cross-reference system
%for figures, tables, equations, etc., is encouraged
%(using \verb"\ref{<name>}" and \verb"\label{<name>}").
%
%\subsection{Endnotes}
%Most \textit{SAGE} journals use endnotes rather than footnotes, so any notes should be coded as \verb+\endnote{<Text>}+.
%Place the command \verb+\theendnotes+ just above the Reference section to typeset the endnotes.
%
%To avoid any confusion for papers that use Vancouver style references,  footnotes/endnotes should be edited into the text.

\section{Copyright statement}
Please  be  aware that the use of  this \LaTeXe\ class file is
governed by the following conditions.

\subsection{Copyright}
Copyright \copyright\ \volumeyear\ SAGE Publications Ltd,
1 Oliver's Yard, 55 City Road, London, EC1Y~1SP, UK. All
rights reserved.

\subsection{Rules of use}
This class file is made available for use by authors who wish to
prepare an article for publication in a \textit{SAGE Publications} journal.
The user may not exploit any
part of the class file commercially.

This class file is provided on an \textit{as is}  basis, without
warranties of any kind, either express or implied, including but
not limited to warranties of title, or implied  warranties of
merchantability or fitness for a particular purpose. There will
be no duty on the author[s] of the software or SAGE Publications Ltd
to correct any errors or defects in the software. Any
statutory  rights you may have remain unaffected by your
acceptance of these rules of use.

\begin{acks}
This class file was developed by Sunrise Setting Ltd,
Brixham, Devon, UK.\\
Website: \url{http://www.sunrise-setting.co.uk}
\end{acks}

\begin{funding}
This research received no specific grant from any funding agency in the public, commercial, or not-for-profit sectors.
\end{funding}

\begin{dci}
	The author declares that there is no conflict of interest.
\end{dci}

\bibliographystyle{SageV}
\bibliography{597_bibliography}

%\begin{thebibliography}{99}
%\bibitem[Kopka and Daly(2003)]{R1}
%Kopka~H and Daly~PW (2003) \textit{A Guide to \LaTeX}, 4th~edn.
%Addison-Wesley.
%
%\bibitem[Lamport(1994)]{R2}
%Lamport~L (1994) \textit{\LaTeX: a Document Preparation System},
%2nd~edn. Addison-Wesley.
%
%\bibitem[Mittelbach and Goossens(2004)]{R3}
%Mittelbach~F and Goossens~M (2004) \textit{The \LaTeX\ Companion},
%2nd~edn. Addison-Wesley.
%
%\end{thebibliography}

\end{document}

### 2021-05-25 python/test.py

@@ -23,6 +23,8 @@ def fit(x):
         area = rs.CurveArea(polyline)[0]
         rs.Redraw()
         return area
+    else:
+        return 0
     # except:
     #     area = 0
     #     # print('oopsies')

### 2021-05-26 writing/essay_latex/draft3.tex
draft 3 of article complete

@@
 
-\documentclass[sagev,times,Review]{sagej}
+\documentclass[sagev,doublespace,times,Review]{sagej}
 % add 'doublespace'
@@
 \usepackage{moreverb,url}
+
+\usepackage{graphicx}
 
@@
 \begin{abstract}
+\textit{***Needs refining and reducing, IJAC word count limit for abstract is 150, and much of my abstract was adapted into my intro- will update with final draft at end of quarter, working findings in***}
+
 The United States is experiencing an unprecedented housing crisis, resulting in a rising population of unhoused peoples and the inability to become a homeowner or to even afford monthly rent in many cities.
@@
 However, a 2019 city survey shows that there are calls for an increased focus on sustainability and cost.
-This research intends to explore whether the use of genetic algorithms via shape grammar methodology to optimize DADU plans to site context, increases building performance or further encourages construction.
+This research intends to explore whether the use of genetic algorithms to optimize DADU windows increases building performance and further encourages construction.
 The proposed methodology begins by reading example site data from the city of Seattle including building and vegetation context and rental/land prices from Seattle GIS and Zillow, respectively.
@@
 Resulting designs will be analyzed and compared via energy performance simulation.
-The end-goal is to develop a computational tool using the aforementioned system to conduct automated site analysis, as well as parametric generation of DADUs.\end{abstract}
+\end{abstract}
 
-\keywords{Genetic Algorithm, ADU, Optimization, Simulation, EUI, Grasshopper, Galapagos}
+\keywords{Genetic algorithm, ADU, optimization, simulation, EUI, grasshopper, galapagos}
 
@@
 
+\section{\textit{TODO:}}
+\begin{itemize}
+	\item Revise and essentially rewrite abstract to fit within journal constraints
+	\item Following points are due to learning LaTeX formatting and how to export Zotero bibliography as a BibLaTeX file to be read and able to cite within the text editor I am using (currently using a work-around): \begin{itemize}
+	\item Add supporting claims into introduction from sources read and researched
+	\item Add supporting claims into methods from sources read and researched \end{itemize}
+	\item Complete explanation of Honeybee energy model and Galapagos genetic solver subsections within methods.
+	\item Add results section (still running the tool on two remaining sites [Madrona and Minor locations])
+	\item Add data visualizations of results, additional simplified plans of DADUs and sites
+	\item Flesh out conclusion, refine portion on future work since pivoting scope and direction of research in week 5
+\end{itemize}
+
 \section{Introduction}
-The United States is amid an unprecedented housing crisis, stemming from issues outside of housing, such as neoliberal cuts to social benefits spending and the increasing privatization of essentially every aspect of American life.
+The United States is amidst an unprecedented housing crisis, stemming from parallel crises and outside issues, such as the neoliberal cuts to social benefits spending and increasing privatization of essentially every aspect of American life.
 The result is a rising population of unhoused peoples and the inability for many to become a homeowner or to even afford minimum monthly rent in many cities.
@@
 
-%TODO Expand background portion of intro
-
 Seattle, Washington; Portland, Oregon; and Vancouver, British Columbia are exploring the use of accessory dwelling units (ADUs) as one method to combat surging housing prices.
@@
 Additionally, ADUs are often designed to be rented, generating supplementary income for the homeowner.
-Ten detached ADU (or DADU) designs are pre-approved by the City of Seattle and are available online to entice homeowners.
-However, a 2019 city survey shows that there are calls for an increased focus on sustainability and cost in the construction of ADUs\cite{seattlePreapprovedPlansAccessory2019}.
-These out of the box designs do not offer the scalability or energy efficiency that an ADU designed specifically per site offers.
+Historically, homeowner’s associations and other local organizations have  been resistant to any proposed density increases through zoning or other means.
+However, Seattle and the other previously mentioned cities in the Pacific Northwest have succeeded in allowing for the construction of ADUs in recent decades, with these groups beginning to see ADUs as a positive.
+ADUs offer many positives beyond an increased housing supply such as a path for homeowners to generate income and accommodating households with multigenerational or caretaking needs.
 
-%TODO Expand reason for why this tool is necessary to drive ADU demand
+Beginning in 2014, the City of Seattle has moved to increase production of ADUs through various resolutions and legislation, removing barriers to entry and highlighting strategies for encouragement.
+In late 2019, a call for detached ADU (or DADU) designs was issued to designers.
+As a result, ten DADU designs were pre-approved by the City of Seattle and are available online to entice homeowners.The pre-approved designs are free, but still require standard permitting fees to be paid for approval and come with one significant downside of many other pre-designed structures- a lack of contextual design and individualization.
+This one-size-fits-all approach can have drastic effects upon energy consumption, reducing the appeal of such an investment.
+Further, these designs do not offer the scalability or energy efficiency that an ADU designed specifically per site offers.
 
-Homeowner’s associations and other local organizations have historically fought back against any proposed density increases through zoning or other method.
-However, Seattle and the other aforementioned cities in the Pacific Northwest have succeeded in allowing for the construction of ADUs in recent decades.
-The pre-approved designs are free, but require payment of around one thousand dollars [check this] for approval and come with one significant downside of many other pre-designed structures- a lack of contextual design and individualization.
+ Coincidentally, a November 2019 city survey shows that there are calls for an increased focus on sustainability and cost in the construction of ADUs\cite{seattlePreapprovedPlansAccessory2019}.
+This survey was positioned before the official call for submissions was broadcast.
+'Low cost' was the top ranked criteria with 48\% of responses listing 'very important', followed by 'green building' being ranked similarly at 46\%.
+Other suggestions sourced from surveyed respondents included low environmental costs, site specific considerations, and predictability.
 
-This research explores whether the use of genetic algorithms via shape grammar methodology can effectively optimize DADU plans to site context, increases building performance or further encourages construction.
-The proposed methodology begins by reading example site data from the city of Seattle including building and vegetation context from the Seattle GIS.
-Next, a genetic algorithm explores the design space for a viable floor plan solution based on a fitness function.
-This fitness function evaluates individual designs according to predefined traits.
-Traits to evaluate include window to wall ratio, insulation depth/type, ventilation strategy, and shading technique.
-Locating the correct combination of traits to minimize (or maximize) which results in a higher performance DADU is the desired outcome.
-Resulting designs will be analyzed and compared via energy performance simulation.
-The end-goal is to develop a computational tool using the aforementioned system to conduct automated site analysis, as well as parametric generation of DADUs.
-
-\section{Literature Review}
-%TODO include feedback info on information from seattle city ADU survey- see obsidian notes: city survey respondents said “»  **Successful designs respond to site and context.** Some respondents lauded design choices that reflect a site’s opportunities and constraints or the architecture and climate of the Pacific Northwest.
-“also: “**Various types of spaces are valuable.** Respondents expressed interest in designs that include patios, rooftop decks, storage, and garages.
-“ - **SPACES TO CONSIDER FOR ALGORITHM** also: **DIY opportunities.** Respondents encouraged designs that allow homeowners to do a substantial amount of the construction themselves.
-“ - **SELF BUILD//NO ARCHITECT/BUILDER INVOLVEMENT don’t say no to this yet!!!
-(family/friend labor?)**also: **Flexibility.** Respondents emphasized the desirability of allowing mix-and-match options, such as multiple roof options, that make the design somewhat customizable.
-“ - **SHAPE GRAMMAR?**
-
+%TODO expand here, but keep flow into methods section
 \section{Methods} 
 \subsection{Design constraints and decisions}
-%TODO code restrictions: height limitations, lot size, etc via aduniverse-seattlecitygis.hub.arcgis.com/pages/code
-%TODO touch on how 3 varying sites were chosen in 3 neighborhoods: central district, first hill and madrona
-%TODO include table of R values for each layer, as well as U value of wall assembly, description of wood frame construction chosen, and why chosen.
+The ability to show users the estimated annual energy consumption of a new construction DADU would have a major impact on driving ADU construction demand based on survey responses noted.
+Using an optimization method such as a genetic algorithm can allow architects to arrive at the most effective design solution on a site specific basis while \textit{predictably} minimizing both environmental and financial costs.
+This research explores the use of a genetic solver to find the optimal design solution of window-to-wall ratios (WWR) for a simple DADU form resulting in the lowest EUI value.
+Three sites in similar neighborhoods in the city of Seattle were selected based on a combination of proximity and slight density variance.
+The tool adheres to land use code to find estimated buildable area for a DADU on each site, creates a simple box-like form representing the DADU and then varies the WWR value for each of the four walls.
+A genetic solver plugin is then allowed to vary these ratios to find an optimal solution using a process largely mimicking natural genetic processes.
+Finally, an energy model is created from each individual iteration to find an estimate for overall EUI.
+From here, the process loops until an optimal solution (in this case, via minimizing EUI) is found.
+
+%TODO merge these two (upper and lower) sections together
+Site locations were chosen through the use of the \textit{ADUniverse} tool, developed by the University of Washington researchers at the eScience Institute\cite{ADUniverseToolEScience}.
+ADUniverse allows for users to investigate the possibility of constructing an ADU through a map-based user interface.
+Chosen sites are located within the Madrona, Minor, and Central District neighborhoods of Seattle, WA.
+All three sites are zoned under SF5000, a single family residential zoning.
+Site data was extracted from the city of Seattle's GIS website, converted into an AutoCAD .DWG file, and then imported into and extruded within Rhinoceros 6 (Rhino), a 3D modeling software.
+Lot lines surrounding each house are fed into the Grasshopper definition, coupled with constraints in line with the current Seattle Land Use Code.
+Such universal constraints for ADUs on lots located within SF5000 zoning are as is shown below \cite{ADURules}:
+
+% itemized list of zoning restrictions for each site
+\begin{itemize}
+	\item Lot size must be at least 3200 square feet
+	\item Minimum lot width of 25 feet
+	\item Minimum lot length of 70 feet
+	\item DADU can occupy \textit{at most} 60\% of rear yard	
+	\item Front yard of 20 feet, side yards of 5 feet, rear yard of 25  feet
+	\item DADU height limit varies based on lot width: \begin{itemize}
+	\item 14 feet for lot widths less than 30 feet
+	\item 16 feet for lot widths between 40 and 50 feet
+	\item 18 feet for lot widths exceeding 50 feet \end{itemize}
+\end{itemize}
+
+These constraints are built into the Grasshopper definition in such a way that the tool only requires lot extends and the outline of the existing residential structure to develop the buildable area for any DADU construction.
+The DADU is represented as a simple box with dimensions of (fill this in once all 3 sites have been finalized).
+This minimal geometry is located congruent to lot lines and the existing structure.
+Under current Seattle Residential Building Code (2018), minimum R-values for wall and ceiling assemblies are R-21 and R-49, respectively.
+Light wood frame construction was chosen for all three sites, as the majority of residential construction in the United States and in Seattle specifically falls under this category due to price and ubiquity.
+For this project, wall assembly was specified to a total R-value of 29.232 (see Table \ref{table:wall-assembly}.).
+\textit{\textbf{***Insert second table or append data to existing table of slab and ceiling R-values***}}
+
+Next, the variable aspect of the DADU lies in the window-to-wall ratios of each of the four facades, which line up with the cardinal directions.
+The ratio between solid and glazed facade areas is allowed to shift between
+
+\begin{table}[]
+\centering
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{lcccc}
+\hline
+\multicolumn{1}{|l|}{\textbf{Assembly}} &
+  \multicolumn{1}{l|}{\textbf{Notes}} &
+  \multicolumn{1}{c|}{\textbf{R-Value}} &
+  \multicolumn{1}{l|}{\textbf{Path 1 (Insulation)}} &
+  \multicolumn{1}{l|}{\textbf{Path 2 (Framing)}} \\ \hline
+\multicolumn{1}{|l|}{Air Film} &
+  \multicolumn{1}{c|}{Interior} &
+  \multicolumn{1}{c|}{0.68} &
+  \multicolumn{1}{c|}{0.68} &
+  \multicolumn{1}{c|}{0.68} \\ \hline
+\multicolumn{1}{|l|}{Gypsum Board} &
+  \multicolumn{1}{c|}{1/2"} &
+  \multicolumn{1}{c|}{0.45} &
+  \multicolumn{1}{c|}{0.45} &
+  \multicolumn{1}{c|}{0.45} \\ \hline
+\multicolumn{1}{|l|}{Vapor Barrier} &
+  \multicolumn{1}{c|}{} &
+  \multicolumn{1}{c|}{0} &
+  \multicolumn{1}{c|}{0} &
+  \multicolumn{1}{c|}{0} \\ \hline
+\multicolumn{1}{|l|}{Fiberglass Insulation} &
+  \multicolumn{1}{c|}{} &
+  \multicolumn{1}{c|}{3.5} &
+  \multicolumn{1}{c|}{28} &
+  \multicolumn{1}{c|}{-} \\ \hline
+\multicolumn{1}{|l|}{2" x 8" Wood Stud} &
+  \multicolumn{1}{c|}{16" o.c.} &
+  \multicolumn{1}{c|}{14.68} &
+  \multicolumn{1}{c|}{-} &
+  \multicolumn{1}{c|}{14.68} \\ \hline
+\multicolumn{1}{|l|}{Air Gap} &
+  \multicolumn{1}{c|}{1"} &
+  \multicolumn{1}{c|}{1} &
+  \multicolumn{1}{c|}{1} &
+  \multicolumn{1}{c|}{1} \\ \hline
+\multicolumn{1}{|l|}{Wood Cladding} &
+  \multicolumn{1}{c|}{3/4"} &
+  \multicolumn{1}{c|}{0.93} &
+  \multicolumn{1}{c|}{0.93} &
+  \multicolumn{1}{c|}{0.93} \\ \hline
+\multicolumn{1}{|l|}{Air Film} &
+  \multicolumn{1}{c|}{Exterior} &
+  \multicolumn{1}{c|}{0.17} &
+  \multicolumn{1}{c|}{0.17} &
+  \multicolumn{1}{c|}{0.17} \\ \hline
+ &
+  \multicolumn{1}{c|}{} &
+  \multicolumn{1}{c|}{\textbf{R-Total:}} &
+  \multicolumn{1}{c|}{31.23} &
+  \multicolumn{1}{c|}{17.91} \\ \cline{3-5} 
+ &
+  \multicolumn{1}{l}{} &
+  \multicolumn{1}{l}{} &
+  \textit{x 85\%} &
+  \textit{x 15\%} \\
+ &
+  \multicolumn{1}{l}{} &
+  \multicolumn{1}{l}{} &
+  26.5455 &
+  2.6865 \\ \cline{3-5} 
+ &
+  \multicolumn{1}{l|}{} &
+  \multicolumn{1}{c|}{\textbf{Wall R-Value:}} &
+  \multicolumn{2}{c|}{29.232} \\ \cline{3-5} 
+ &
+  \multicolumn{1}{l|}{} &
+  \multicolumn{1}{l|}{\textbf{Wall U-Value:}} &
+  \multicolumn{2}{c|}{0.0342} \\ \cline{3-5} 
+ &
+  \multicolumn{1}{l}{} &
+  \multicolumn{1}{l}{} &
+  \multicolumn{1}{l}{} &
+  \multicolumn{1}{l}{}
+\end{tabular}%
+}
+\caption{Wall Assembly R- and U-Values}
+\label{table:wall-assembly}
+\end{table}
+
 \subsection{Energy simulation and analysis}
+Honeybee, part of the open-source Ladybug Tools plugin suite for Grasshopper and Rhino, has become an industry standard for daylighting simulation and energy modeling within architecture and engineering.
+%TODO expand how Honeybee works with energyplus and other tools to calculate energy statistics.
+Trees are represented via spherical meshes created using Grasshopper, a visual programming tool for Rhinoceros, and placed where necessary based on the .DWG file.
+These trees vary in opaque triangulated mesh coverage to simulate foliage cover and to more accurately cast shadows.
+Originally, trees were represented as simple cylindrical solids, which increased optimization rate and speed of energy analysis, but decreased validity and accuracy of results.
+
 \subsection{Genetic algorithm}
+Genetic algorithms (GA) operate using logic that largely mimics evolutionary biological models to optimize a value.
+Fitness functions are the means by which to objectively evaluate the strength of individual designs.
+In this case, the fitness function specified works to minimize the resultant EUI value of each DADU design.
+Work was initially done using a purely Python-based GA script, but later pivoted to utilizing Galapagos, a genetic solver plugin that ships with Grasshopper.
+Galapagos offers a simple user interface and more direct integration within Grasshopper at the cost of advanced options.
+These options however were not necessary for the level of optimization required within this project.
 
+%Together, the contextual 3D model, the specific design iteration of the DADU with varying WWRs, and the EUI value 
 \section{Results}
@@
 \section{Conclusion}
-
-\subsection{Future work}
-Benefit of a pure python tool 
+%TODO Add main conclusion section
+%TODO Benefit of a pure python tool 
 Looking forward, moving on from the Grasshopper platform gives the benefit of non-reliance on developers to maintain the software in which the tool depends.
@@
 Additionally, Python is used within Rhino/Grasshopper (or a flavor thereof), provides many useful math and geometry libraries, and has options for injection into a web app (Flask and Django).
-Limitations of Galapagos/Grasshopper workflow
-Utilizing Galapagos and Grasshopper offers a ‘sandbox’ environment to begin to understand genetic algorithms and to arrive at tangible design solutions faster than using a home-brewed algorithm.
+%TODO Limitations of Galapagos/Grasshopper workflow
+Utilizing Grasshopper and Galapagos offers a ‘sandbox’ environment to begin to understand genetic algorithms and to arrive at tangible design solutions faster than using a home-brewed algorithm.
 Galapagos offers many fewer input parameters and a much narrower scope in which to define the fitness function.
@@
 
-Figures are called in as follows:
-\begin{verbatim}
-\begin{figure}
-\centering
-\includegraphics{<figure name>}
-\caption{<Figure caption>}
-\end{figure}
-\end{verbatim}
+***User ability to override window sizing and placement- feedback loop?
 
+%Figures are called in as follows:
+%\begin{verbatim}
 %\begin{figure}
-%\setlength{\fboxsep}{0pt}%
-%\setlength{\fboxrule}{0pt}%
-%\begin{center}
-%\begin{boxedverbatim}
-%\begin{table}
-%\small\sf\centering
-%\caption{<Table caption.>}
-%\begin{tabular}{<table alignment>}
-%\toprule
-%<column headings>\\
-%\midrule
-%<table entries
-%(separated by & as usual)>\\
-%<table entries>\\
-%.
-%.
-%.\\
-%\bottomrule
-%\end{tabular}
-%\end{table}
-%\end{boxedverbatim}
-%\end{center}
-%\caption{Example table layout.\label{F2}}
+%\centering
+%\includegraphics{<figure name>}
+%\caption{<Figure caption>}
 %\end{figure}
+%\end{verbatim}
 
@@
 \begin{acks}
+%TODO acknowledge Tomas, Teresa, Rick Mohler, eScience for Social Good team behind ADUniverse, Alex, other students in the class who reviewed my shit.
 This class file was developed by Sunrise Setting Ltd,

### 2021-05-28 writing/essay_latex/draft3.tex
draft3 complete

@@
 
-\subsection{Energy simulation and analysis}
+\subsection{Energy model and analysis}
 Honeybee, part of the open-source Ladybug Tools plugin suite for Grasshopper and Rhino, has become an industry standard for daylighting simulation and energy modeling within architecture and engineering.

### 2021-06-08 writing/essay_latex/draft3.tex
draft edits

@@
 
-\title{Energy simulation and analysis of accessory dwelling unit: 
+\title{Energy simulation and analysis of accessory dwelling units: 
 an evolutionary approach}
@@
 
-The United States is experiencing an unprecedented housing crisis, resulting in a rising population of unhoused peoples and the inability to become a homeowner or to even afford monthly rent in many cities.
-Solutions to this dilemma are neither straightforward nor definite.
-Seattle, Washington; Portland, Oregon; and Vancouver, British Columbia are exploring the use of accessory dwelling units (ADUs) as one method to combat surging housing prices.
-Use of ADUs is an effective means of increasing housing density without replacing single family housing zones with new multi family residential construction.
-Additionally, ADUs are often designed to be rented, generating supplementary income for the homeowner.
 Ten detached ADU (or DADU) designs are pre-approved by the City of Seattle and are freely available online to entice homeowners.
@@
 \section{Introduction}
-The United States is amidst an unprecedented housing crisis, stemming from parallel crises and outside issues, such as the neoliberal cuts to social benefits spending and increasing privatization of essentially every aspect of American life.
+The United States is amidst an unprecedented housing crisis, stemming from parallel crises, such as the neoliberal cuts to social benefits spending and increasing privatization of essentially every aspect of American life.
 The result is a rising population of unhoused peoples and the inability for many to become a homeowner or to even afford minimum monthly rent in many cities.
 Solutions to this dilemma are neither straightforward nor definite.
-However, as designers and architects, this problem can only be addressed at the symptom-level, while advocating at the root cause.
+However, designers and architects can only address the symptom-level problems, while advocating towards change directed at the root cause.
 
@@
 
-These constraints are built into the Grasshopper definition in such a way that the tool only requires lot extends and the outline of the existing residential structure to develop the buildable area for any DADU construction.
+These constraints are built into the Grasshopper definition in such a way that the tool only requires lot extents and the outline of the existing residential structure to develop the buildable area for any DADU construction.
 The DADU is represented as a simple box with dimensions of (fill this in once all 3 sites have been finalized).
@@
 %TODO expand how Honeybee works with energyplus and other tools to calculate energy statistics.
+%TODO change information on how trees were represented.
 Trees are represented via spherical meshes created using Grasshopper, a visual programming tool for Rhinoceros, and placed where necessary based on the .DWG file.
 These trees vary in opaque triangulated mesh coverage to simulate foliage cover and to more accurately cast shadows.
-Originally, trees were represented as simple cylindrical solids, which increased optimization rate and speed of energy analysis, but decreased validity and accuracy of results.
+In initial tests, trees were represented as simple cylindrical solids, which increased optimization rate and speed of energy analysis, but decreased validity and accuracy of results.
 

### 2021-06-10 writing/essay_latex/draft3.tex
writing progress

@@
 
-\documentclass[sagev,doublespace,times,Review]{sagej}
+\documentclass[sagev,times,Royal]{sagej}
 % add 'doublespace'
@@
 
-\author{Preston Pape\affilnum{1}}
+\author{P. Arthur Pape\affilnum{1}}
 
@@
 
-%\corrauth{}
+\corrauth{P. Arthur Pape}
 
@@
 \begin{abstract}
-\textit{***Needs refining and reducing, IJAC word count limit for abstract is 150, and much of my abstract was adapted into my intro- will update with final draft at end of quarter, working findings in***}
-
-Ten detached ADU (or DADU) designs are pre-approved by the City of Seattle and are freely available online to entice homeowners.
-However, a 2019 city survey shows that there are calls for an increased focus on sustainability and cost.
-This research intends to explore whether the use of genetic algorithms to optimize DADU windows increases building performance and further encourages construction.
-The proposed methodology begins by reading example site data from the city of Seattle including building and vegetation context and rental/land prices from Seattle GIS and Zillow, respectively.
-Next, a genetic algorithm explores the design space for a viable floor plan solution based on a fitness function.
-This fitness function evaluates individual designs according to predefined traits.
-Traits to evaluate include window to wall ratio, insulation depth/type, ventilation strategy, and shading technique.
-Locating the correct combination of traits to minimize (or maximize) which results in a higher performance DADU is the desired outcome.
-Resulting designs will be analyzed and compared via energy performance simulation.
+Accessory dwelling units (ADUs) are surging in popularity as cities push for their construction to aid in combatting the housing crisis.
+In Seattle, recent survey data calls for increased focus on sustainability and cost.
+Lowering energy use intensity (EUI) of ADUs allows for the targeting of both desires which homeowners hold.
+The effect of varying window-to-wall ratios (WWRs) on EUI is simulated using EnergyPlus and Honeybee.
+Galapagos, a genetic solver plugin for Grasshopper is used to vary window-to-wall ratio and optimize ADU energy use intensity on three single family residential lots in Seattle, Washington.
+Site context is constructed using OpenStreetMap GIS data and zoning restrictions are incorporated into the Grasshopper definition.
+Future work is planned to utilize output data from this tool as the training dataset for a machine learning model.
+Accurate machine learning prediction of energy use intensity instead of extended simulation would increase pragmatic use of such tools.
 \end{abstract}
 
-\keywords{Genetic algorithm, ADU, optimization, simulation, EUI, grasshopper, galapagos}
+\keywords{Genetic algorithm, accessory dwelling unit, housing, optimization, simulation, energy use intensity, grasshopper, galapagos}
 
@@
 
-\section{\textit{TODO:}}
-\begin{itemize}
-	\item Revise and essentially rewrite abstract to fit within journal constraints
-	\item Following points are due to learning LaTeX formatting and how to export Zotero bibliography as a BibLaTeX file to be read and able to cite within the text editor I am using (currently using a work-around): \begin{itemize}
-	\item Add supporting claims into introduction from sources read and researched
-	\item Add supporting claims into methods from sources read and researched \end{itemize}
-	\item Complete explanation of Honeybee energy model and Galapagos genetic solver subsections within methods.
-	\item Add results section (still running the tool on two remaining sites [Madrona and Minor locations])
-	\item Add data visualizations of results, additional simplified plans of DADUs and sites
-	\item Flesh out conclusion, refine portion on future work since pivoting scope and direction of research in week 5
-\end{itemize}
+%TODO Complete explanation of Honeybee energy model and Galapagos genetic solver subsections within methods.
+%TODO Add data visualizations of results, additional simplified plans or renderings of DADUs and sites
+%TODO Flesh out conclusion, limit and refine portion on future work towards machine learning surrogate modeling
 
 \section{Introduction}
-The United States is amidst an unprecedented housing crisis, stemming from parallel crises, such as the neoliberal cuts to social benefits spending and increasing privatization of essentially every aspect of American life.
+The United States is amidst an unprecedented housing crisis, stemming from parallel crises: from neoliberal austerity spending cuts towards social benefits to the near-virtual lack of social housing and rent control programs in the United States to the ever-increasing privatization of essentially every aspect of American life\cite{10.2307/26297969}.
 The result is a rising population of unhoused peoples and the inability for many to become a homeowner or to even afford minimum monthly rent in many cities.
@@
 Additionally, ADUs are often designed to be rented, generating supplementary income for the homeowner.
-Historically, homeowner’s associations and other local organizations have  been resistant to any proposed density increases through zoning or other means.
+Historically, homeowner’s associations and other local organizations have  been resistant to any proposed density increases through zoning or other means\cite{10.2307/24392672}.
 However, Seattle and the other previously mentioned cities in the Pacific Northwest have succeeded in allowing for the construction of ADUs in recent decades, with these groups beginning to see ADUs as a positive.
@@
 
-Beginning in 2014, the City of Seattle has moved to increase production of ADUs through various resolutions and legislation, removing barriers to entry and highlighting strategies for encouragement.
+Beginning in 2014, the City of Seattle has moved to increase production of ADUs through various resolutions and legislation, in turn removing barriers to entry and highlighting strategies for encouragement.
 In late 2019, a call for detached ADU (or DADU) designs was issued to designers.
-As a result, ten DADU designs were pre-approved by the City of Seattle and are available online to entice homeowners.The pre-approved designs are free, but still require standard permitting fees to be paid for approval and come with one significant downside of many other pre-designed structures- a lack of contextual design and individualization.
+As a result, ten DADU designs were pre-approved by the City of Seattle and are available online to entice homeowners\cite{ADUniversePreapprovedADU}.
+The pre-approved designs are free, but still require standard permitting fees to be paid for approval and come with one significant downside of many other pre-designed structures: a lack of contextual design and individualization.
 This one-size-fits-all approach can have drastic effects upon energy consumption, reducing the appeal of such an investment.
@@
 Other suggestions sourced from surveyed respondents included low environmental costs, site specific considerations, and predictability.
+The ability to show users the estimated annual energy consumption of a new construction DADU would have a major impact on driving construction demand based on survey responses noted.
+Using an optimization method such as a genetic algorithm can allow architects to arrive at the most effective design solution on a site specific basis while predictably minimizing both environmental and financial costs.
 
-%TODO expand here, but keep flow into methods section
 \section{Methods} 
 \subsection{Design constraints and decisions}
-The ability to show users the estimated annual energy consumption of a new construction DADU would have a major impact on driving ADU construction demand based on survey responses noted.
-Using an optimization method such as a genetic algorithm can allow architects to arrive at the most effective design solution on a site specific basis while \textit{predictably} minimizing both environmental and financial costs.
 This research explores the use of a genetic solver to find the optimal design solution of window-to-wall ratios (WWR) for a simple DADU form resulting in the lowest EUI value.
 Three sites in similar neighborhoods in the city of Seattle were selected based on a combination of proximity and slight density variance.
-The tool adheres to land use code to find estimated buildable area for a DADU on each site, creates a simple box-like form representing the DADU and then varies the WWR value for each of the four walls.
+The tool adheres to Seattle's land use code to find estimated buildable area for a DADU on each site, creates a simple box-like form representing the DADU and then varies the WWR value for each of the four walls.
 A genetic solver plugin is then allowed to vary these ratios to find an optimal solution using a process largely mimicking natural genetic processes.
@@
 Lot lines surrounding each house are fed into the Grasshopper definition, coupled with constraints in line with the current Seattle Land Use Code.
-Such universal constraints for ADUs on lots located within SF5000 zoning are as is shown below \cite{ADURules}:
+Such universal constraints for ADUs on lots located within SF5000 zoning are as follows\cite{ADURules}:
 
@@
 
+%TODO revise assembly table to match assembly used in grasshopper
 \begin{table}[]
@@
 Honeybee, part of the open-source Ladybug Tools plugin suite for Grasshopper and Rhino, has become an industry standard for daylighting simulation and energy modeling within architecture and engineering.
-%TODO expand how Honeybee works with energyplus and other tools to calculate energy statistics.
-%TODO change information on how trees were represented.
-Trees are represented via spherical meshes created using Grasshopper, a visual programming tool for Rhinoceros, and placed where necessary based on the .DWG file.
-These trees vary in opaque triangulated mesh coverage to simulate foliage cover and to more accurately cast shadows.
-In initial tests, trees were represented as simple cylindrical solids, which increased optimization rate and speed of energy analysis, but decreased validity and accuracy of results.
+Honeybee acts as a bridge between modeled Rhino geometry and the United States Department of Energy's EnergyPlus simulation tool.
+Modeled geometry is constructed as rooms and zones within Honeybee, offering greater granularity.
+Using four zones and a core allows Honeybee and EnergyPlus to take into account the effect of glazing proportions on each facade has on heating and cooling EUI in more detail.
+EnergyPlus operates using various abstractions and simplifications and these are important to note.
+Each zone and surface is only able to hold one temperature value at a time- each singular value is averaged across the zone or surface.
+Each DADU simulated is represented as four zones and a core, the standard for simulations of this type.
+Alternatively, using one zone instead would have assisted with quicker simulation times, and therefore quicker optimization speed.
+
+Similarly, trees are represented as simple cylindrical solids, in order to keep energy simulation speed quick and to reduce the time necessary to reach an optimal solution.
+In initial tests, trees were represented as spherical meshes created using Grasshopper.
+These more-detailed trees varied in opaque triangulated mesh coverage to simulate foliage cover and to more accurately cast shadows.
+It is important to note that with access to greater computing power it could become reasonable to represent and simulate trees in the original format.
+
+%TODO all inputs, geometry or otherwise required for honeybee/energyplus
 
 \subsection{Genetic algorithm}
-Genetic algorithms (GA) operate using logic that largely mimics evolutionary biological models to optimize a value.
+Genetic algorithms (GA) operate using logic that mimics biological evolution to optimize a value.
 Fitness functions are the means by which to objectively evaluate the strength of individual designs.
@@
 %TODO Limitations of Galapagos/Grasshopper workflow
+%TODO ***User ability to override window sizing and placement- feedback loop?
 Utilizing Grasshopper and Galapagos offers a ‘sandbox’ environment to begin to understand genetic algorithms and to arrive at tangible design solutions faster than using a home-brewed algorithm.
@@
 In its out-of-the-box form, Galapagos only accepts number sliders as input and can only optimize integers and floats- in reality, there is not means in which to define a true fitness function, only numerical values to target.
-
-***User ability to override window sizing and placement- feedback loop?
+%TODO future work of training machine learning model using energyplus data to replace simulation timesink
 
@@
 
-\section{Copyright statement}
-Please  be  aware that the use of  this \LaTeXe\ class file is
-governed by the following conditions.
-
-\subsection{Copyright}
-Copyright \copyright\ \volumeyear\ SAGE Publications Ltd,
-1 Oliver's Yard, 55 City Road, London, EC1Y~1SP, UK. All
-rights reserved.
-
-\subsection{Rules of use}
-This class file is made available for use by authors who wish to
-prepare an article for publication in a \textit{SAGE Publications} journal.
-The user may not exploit any
-part of the class file commercially.
-
-This class file is provided on an \textit{as is}  basis, without
-warranties of any kind, either express or implied, including but
-not limited to warranties of title, or implied  warranties of
-merchantability or fitness for a particular purpose. There will
-be no duty on the author[s] of the software or SAGE Publications Ltd
-to correct any errors or defects in the software. Any
-statutory  rights you may have remain unaffected by your
-acceptance of these rules of use.
-
 \begin{acks}
-%TODO acknowledge Tomas, Teresa, Rick Mohler, eScience for Social Good team behind ADUniverse, Alex, other students in the class who reviewed my shit.
-This class file was developed by Sunrise Setting Ltd,
-Brixham, Devon, UK.\\
-Website: \url{http://www.sunrise-setting.co.uk}
+The author would like to thank Professor Tomas Mendez Echenagucia for his genetic algorithm expertise and input; Teresa Moroseos for her assistance with Honeybee debugging; Professor Alex Anderson and fellow students in Arch597 at University of Washington for feedback and motivation.
 \end{acks}
@@
 \begin{funding}
-This research received no specific grant from any funding agency in the public, commercial, or not-for-profit sectors.
+The author(s) received no financial support for the research, authorship, and/or publication of this article.
 \end{funding}
@@
 \begin{dci}
-	The author declares that there is no conflict of interest.
+	The author(s) declared no potential conflicts of interest with respect to the research, authorship, and/or publication of this article.
 \end{dci}
@@
 
-%\begin{thebibliography}{99}
-%\bibitem[Kopka and Daly(2003)]{R1}
-%Kopka~H and Daly~PW (2003) \textit{A Guide to \LaTeX}, 4th~edn.
-%Addison-Wesley.
-%
-%\bibitem[Lamport(1994)]{R2}
-%Lamport~L (1994) \textit{\LaTeX: a Document Preparation System},
-%2nd~edn. Addison-Wesley.
-%
-%\bibitem[Mittelbach and Goossens(2004)]{R3}
-%Mittelbach~F and Goossens~M (2004) \textit{The \LaTeX\ Companion},
-%2nd~edn. Addison-Wesley.
-%
-%\end{thebibliography}
-
 \end{document}

### 2021-06-11 writing/essay_latex/Energy_simulation_and_analysis_ADU.tex
filename fix

--- 2021-06-10 writing/essay_latex/draft3.tex
@@
 % sage_latex_guidelines.tex V1.20, 14 January 2017
+
+% 4000-6000 words, 2000 here as of everything but results and conclusion complete
 
@@
 Accessory dwelling units (ADUs) are surging in popularity as cities push for their construction to aid in combatting the housing crisis.
-In Seattle, recent survey data calls for increased focus on sustainability and cost.
+In Seattle, Washington, recent survey data calls for increased focus on sustainability and cost.
 Lowering energy use intensity (EUI) of ADUs allows for the targeting of both desires which homeowners hold.
 The effect of varying window-to-wall ratios (WWRs) on EUI is simulated using EnergyPlus and Honeybee.
-Galapagos, a genetic solver plugin for Grasshopper is used to vary window-to-wall ratio and optimize ADU energy use intensity on three single family residential lots in Seattle, Washington.
+Galapagos, a genetic solver plugin for Grasshopper is used to vary window-to-wall ratio and optimize ADU energy use intensity on three single family residential lots in Seattle.
 Site context is constructed using OpenStreetMap GIS data and zoning restrictions are incorporated into the Grasshopper definition.
@@
 \maketitle
-
-%TODO Complete explanation of Honeybee energy model and Galapagos genetic solver subsections within methods.
-%TODO Add data visualizations of results, additional simplified plans or renderings of DADUs and sites
-%TODO Flesh out conclusion, limit and refine portion on future work towards machine learning surrogate modeling
 
@@
 
-\section{Methods} 
-\subsection{Design constraints and decisions}
-This research explores the use of a genetic solver to find the optimal design solution of window-to-wall ratios (WWR) for a simple DADU form resulting in the lowest EUI value.
-Three sites in similar neighborhoods in the city of Seattle were selected based on a combination of proximity and slight density variance.
-The tool adheres to Seattle's land use code to find estimated buildable area for a DADU on each site, creates a simple box-like form representing the DADU and then varies the WWR value for each of the four walls.
+\section{Methodology} 
+This research explores the use of a genetic solver to find the optimal design solution of window-to-wall ratios (WWR) for a simple DADU form resulting in the lowest energy use intensity (EUI) value.
+Three sites in comparable neighborhoods in the city of Seattle were selected based on a combination of proximity and slight density variance.
+The tool adheres to zoning regulations to find estimated buildable area for a DADU on each site, creates a simple box-like form representing the DADU and then varies the WWR value for each of the four walls.
 A genetic solver plugin is then allowed to vary these ratios to find an optimal solution using a process largely mimicking natural genetic processes.
 Finally, an energy model is created from each individual iteration to find an estimate for overall EUI.
-From here, the process loops until an optimal solution (in this case, via minimizing EUI) is found.
+From here, the process loops until an optimal solution is found.
 
-%TODO merge these two (upper and lower) sections together
-Site locations were chosen through the use of the \textit{ADUniverse} tool, developed by the University of Washington researchers at the eScience Institute\cite{ADUniverseToolEScience}.
+Site locations were chosen through the use of the ADUniverse tool, developed by University of Washington researchers at the eScience Institute's \textit{Data Science for Social Good} program\cite{ADUniverseToolEScience}.
 ADUniverse allows for users to investigate the possibility of constructing an ADU through a map-based user interface.
@@
 All three sites are zoned under SF5000, a single family residential zoning.
-Site data was extracted from the city of Seattle's GIS website, converted into an AutoCAD .DWG file, and then imported into and extruded within Rhinoceros 6 (Rhino), a 3D modeling software.
-Lot lines surrounding each house are fed into the Grasshopper definition, coupled with constraints in line with the current Seattle Land Use Code.
-Such universal constraints for ADUs on lots located within SF5000 zoning are as follows\cite{ADURules}:
+Site data was extracted from open source map website OpenStreetMap.org and converted into an AutoCAD .DWG file.
+Next, this 2D linework was imported into and extruded into three dimensions within Rhinoceros 6 (Rhino) and Grasshopper, a 3D modeling software and visual scripting tool for Rhino, respectively.
+Property lines (pulled from the Seattle City GIS tool) for each chosen residential site are fed into the Grasshopper definition, coupled with constraints in-line with the current land use code of Seattle \cite{SeattleLandUse2021}.
+Such universal constraints for ADUs on lots located within SF5000 zoning are as follows\cite{ADUniverseADURules}:
 
-% itemized list of zoning restrictions for each site
+% Itemized list of zoning restrictions for each site
 \begin{itemize}
@@
 	\item 18 feet for lot widths exceeding 50 feet \end{itemize}
+\label{zoning-restrictions}
 \end{itemize}
 
-These constraints are built into the Grasshopper definition in such a way that the tool only requires lot extents and the outline of the existing residential structure to develop the buildable area for any DADU construction.
-The DADU is represented as a simple box with dimensions of (fill this in once all 3 sites have been finalized).
-This minimal geometry is located congruent to lot lines and the existing structure.
+These constraints are built into the Grasshopper definition in such a way that the tool requires simply lot extents and the footprint of the existing residential structure to develop the buildable area for any DADU construction.
+Each DADU is represented as a simple 12' shoebox model with facades oriented  perpendicular to the cardinal directions.
 Under current Seattle Residential Building Code (2018), minimum R-values for wall and ceiling assemblies are R-21 and R-49, respectively.
 Light wood frame construction was chosen for all three sites, as the majority of residential construction in the United States and in Seattle specifically falls under this category due to price and ubiquity.
-For this project, wall assembly was specified to a total R-value of 29.232 (see Table \ref{table:wall-assembly}.).
-\textit{\textbf{***Insert second table or append data to existing table of slab and ceiling R-values***}}
+For this project, total R-values were set to 26.86, 54.592, and 5.73, for wall, roof, and floor assemblies, respectively.
+Wall and roof assemblies are detailed in Tables \ref{table:wall-assembly} and \ref{table:roof-assembly}.
+Floor condition is a standard slab-on-grade with rigid insulation below an 8" concrete slab, finished with carpeting.
 
-Next, the variable aspect of the DADU lies in the window-to-wall ratios of each of the four facades, which line up with the cardinal directions.
-The ratio between solid and glazed facade areas is allowed to shift between
-
-%TODO revise assembly table to match assembly used in grasshopper
+% Wall Assembly
 \begin{table}[]
@@
   \multicolumn{1}{c|}{0} \\ \hline
-\multicolumn{1}{|l|}{Fiberglass Insulation} &
-  \multicolumn{1}{c|}{} &
-  \multicolumn{1}{c|}{3.5} &
-  \multicolumn{1}{c|}{28} &
+\multicolumn{1}{|l|}{Rigid Insulation} &
+  \multicolumn{1}{c|}{4"} &
+  \multicolumn{1}{c|}{6} &
+  \multicolumn{1}{c|}{24} &
   \multicolumn{1}{c|}{-} \\ \hline
-\multicolumn{1}{|l|}{2" x 8" Wood Stud} &
+\multicolumn{1}{|l|}{2" x 6" Wood Stud} &
   \multicolumn{1}{c|}{16" o.c.} &
-  \multicolumn{1}{c|}{14.68} &
+  \multicolumn{1}{c|}{12.31} &
   \multicolumn{1}{c|}{-} &
-  \multicolumn{1}{c|}{14.68} \\ \hline
+  \multicolumn{1}{c|}{12.31} \\ \hline
 \multicolumn{1}{|l|}{Air Gap} &
   \multicolumn{1}{c|}{1"} &
-  \multicolumn{1}{c|}{1} &
-  \multicolumn{1}{c|}{1} &
-  \multicolumn{1}{c|}{1} \\ \hline
+  \multicolumn{1}{c|}{2.38} &
+  \multicolumn{1}{c|}{2.38} &
+  \multicolumn{1}{c|}{2.38} \\ \hline
 \multicolumn{1}{|l|}{Wood Cladding} &
@@
   \multicolumn{1}{c|}{\textbf{R-Total:}} &
-  \multicolumn{1}{c|}{31.23} &
-  \multicolumn{1}{c|}{17.91} \\ \cline{3-5} 
+  \multicolumn{1}{c|}{28.61} &
+  \multicolumn{1}{c|}{16.92} \\ \cline{3-5} 
  &
@@
   \multicolumn{1}{l}{} &
-  26.5455 &
-  2.6865 \\ \cline{3-5} 
+  24.3185 &
+  2.538 \\ \cline{3-5} 
  &
@@
   \multicolumn{1}{c|}{\textbf{Wall R-Value:}} &
-  \multicolumn{2}{c|}{29.232} \\ \cline{3-5} 
+  \multicolumn{2}{c|}{26.86} \\ \cline{3-5} 
  &
@@
   \multicolumn{1}{l|}{\textbf{Wall U-Value:}} &
-  \multicolumn{2}{c|}{0.0342} \\ \cline{3-5} 
+  \multicolumn{2}{c|}{0.0372} \\ \cline{3-5} 
  &
@@
 }
-\caption{Wall Assembly R- and U-Values}
+\caption{Wall Assembly}
 \label{table:wall-assembly}
@@
 
+% Roof Assembly
+\begin{table}[]
+\centering
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{lcccc}
+\hline
+\multicolumn{1}{|l|}{\textbf{Assembly}} &
+  \multicolumn{1}{l|}{\textbf{Notes}} &
+  \multicolumn{1}{c|}{\textbf{R-Value}} &
+  \multicolumn{1}{l|}{\textbf{Path 1 (Insulation)}} &
+  \multicolumn{1}{l|}{\textbf{Path 2 (Framing)}} \\ \hline
+\multicolumn{1}{|l|}{Air Film} &
+  \multicolumn{1}{c|}{Interior} &
+  \multicolumn{1}{c|}{0.61} &
+  \multicolumn{1}{c|}{0.61} &
+  \multicolumn{1}{c|}{0.61} \\ \hline
+\multicolumn{1}{|l|}{Gypsum Board} &
+  \multicolumn{1}{c|}{1/2"} &
+  \multicolumn{1}{c|}{0.45} &
+  \multicolumn{1}{c|}{0.45} &
+  \multicolumn{1}{c|}{0.45} \\ \hline
+\multicolumn{1}{|l|}{Vapor Barrier} &
+  \multicolumn{1}{c|}{} &
+  \multicolumn{1}{c|}{0} &
+  \multicolumn{1}{c|}{0} &
+  \multicolumn{1}{c|}{0} \\ \hline
+\multicolumn{1}{|l|}{Cellulose Insulation} &
+  \multicolumn{1}{c|}{Blow-in} &
+  \multicolumn{1}{c|}{47} &
+  \multicolumn{1}{c|}{47} &
+  \multicolumn{1}{c|}{-} \\ \hline
+\multicolumn{1}{|l|}{2" x 10" Wood Rafters} &
+  \multicolumn{1}{c|}{24" o.c.} &
+  \multicolumn{1}{c|}{14.68} &
+  \multicolumn{1}{c|}{-} &
+  \multicolumn{1}{c|}{14.68} \\ \hline
+\multicolumn{1}{|l|}{Rigid Insulation} &
+  \multicolumn{1}{c|}{2"} &
+  \multicolumn{1}{c|}{10} &
+  \multicolumn{1}{c|}{10} &
+  \multicolumn{1}{c|}{10} \\ \hline
+\multicolumn{1}{|l|}{OSB Sheathing} &
+  \multicolumn{1}{c|}{7/16"} &
+  \multicolumn{1}{c|}{0.77} &
+  \multicolumn{1}{c|}{0.77} &
+  \multicolumn{1}{c|}{0.77} \\ \hline
+\multicolumn{1}{|l|}{Asphalt Shingles} &
+  \multicolumn{1}{c|}{} &
+  \multicolumn{1}{c|}{0.44} &
+  \multicolumn{1}{c|}{0.44} &
+  \multicolumn{1}{c|}{0.44} \\ \hline
+\multicolumn{1}{|l|}{Air Film} &
+  \multicolumn{1}{c|}{Exterior} &
+  \multicolumn{1}{c|}{0.17} &
+  \multicolumn{1}{c|}{0.17} &
+  \multicolumn{1}{c|}{0.17} \\ \hline
+ &
+  \multicolumn{1}{c|}{} &
+  \multicolumn{1}{c|}{\textbf{R-Total:}} &
+  \multicolumn{1}{c|}{59.44} &
+  \multicolumn{1}{c|}{27.12} \\ \cline{3-5} 
+ &
+  \multicolumn{1}{l}{} &
+  \multicolumn{1}{l}{} &
+  \textit{x 85\%} &
+  \textit{x 15\%} \\
+ &
+  \multicolumn{1}{l}{} &
+  \multicolumn{1}{l}{} &
+  50.524 &
+  4.068 \\ \cline{3-5} 
+ &
+  \multicolumn{1}{l|}{} &
+  \multicolumn{1}{c|}{\textbf{Wall R-Value:}} &
+  \multicolumn{2}{c|}{54.592} \\ \cline{3-5} 
+ &
+  \multicolumn{1}{l|}{} &
+  \multicolumn{1}{l|}{\textbf{Wall U-Value:}} &
+  \multicolumn{2}{c|}{0.0183} \\ \cline{3-5} 
+ &
+  \multicolumn{1}{l}{} &
+  \multicolumn{1}{l}{} &
+  \multicolumn{1}{l}{} &
+  \multicolumn{1}{l}{}
+\end{tabular}%
+}
+\caption{Roof Assembly}
+\label{table:roof-assembly}
+\end{table}
+
 \subsection{Energy model and analysis}
-Honeybee, part of the open-source Ladybug Tools plugin suite for Grasshopper and Rhino, has become an industry standard for daylighting simulation and energy modeling within architecture and engineering.
+Honeybee, part of the open-source Ladybug Tools plugin suite for Grasshopper and Rhino, has become an industry standard for daylighting simulation and energy modeling within architecture and engineering, largely due to its open source nature\cite{mackey2018tool}.
 Honeybee acts as a bridge between modeled Rhino geometry and the United States Department of Energy's EnergyPlus simulation tool.
 Modeled geometry is constructed as rooms and zones within Honeybee, offering greater granularity.
-Using four zones and a core allows Honeybee and EnergyPlus to take into account the effect of glazing proportions on each facade has on heating and cooling EUI in more detail.
-EnergyPlus operates using various abstractions and simplifications and these are important to note.
+EnergyPlus operates using various abstractions and simplifications and these are important to note\cite{EnergyPlusTMVersionDocumentation2021}.
 Each zone and surface is only able to hold one temperature value at a time- each singular value is averaged across the zone or surface.
-Each DADU simulated is represented as four zones and a core, the standard for simulations of this type.
-Alternatively, using one zone instead would have assisted with quicker simulation times, and therefore quicker optimization speed.
+Each DADU simulated is represented as a single zone spanning the entire structure.
+Alternatively, using four zones and a core instead would have increased resolution of results, but at the cost of much longer simulation times.
+For building typologies larger than ADUs, this discrepancy between EUI simulation results using one or five (four and a core) zones has been studied\cite{SingleZoneVs2018}.
+Results showed EUI estimates experienced 4.23\% variance on average for cities tested.
+Further, for Seattle specifically, results varied by only 2.12\%.
 
@@
 It is important to note that with access to greater computing power it could become reasonable to represent and simulate trees in the original format.
+Together with other context, including the associated residence and nearby residences capable of projecting shadows onto the DADU, the tree models were assigned as shades in the Honeybee model.
 
-%TODO all inputs, geometry or otherwise required for honeybee/energyplus
+The final input required for the Honeybee/EnergyPlus model is a weather file, in .EPW (EnergyPlus Weather File) format.
+These files contain yearly climate data for a geographic location; in this case data was collected at the King County International Airport - Boeing Field in Seattle, Washington.
+The American Society of Heating, Refrigerating, and Air-Conditioning Engineers (ASHRAE) defines climate zones for use with EnergyPlus and other tools/calculations.
+Seattle falls within the ASHRAE 169-2006 standard 4C climate, defined as \textit{Mixed - Marine}.
+This climate type is utilized within the Grasshopper script to define the wood frame construction type for the Seattle climate.
 
@@
 Fitness functions are the means by which to objectively evaluate the strength of individual designs.
-In this case, the fitness function specified works to minimize the resultant EUI value of each DADU design.
+In this case, the fitness value minimized is the resultant EUI value corresponding to each set of window-to-wall ratios.
 Work was initially done using a purely Python-based GA script, but later pivoted to utilizing Galapagos, a genetic solver plugin that ships with Grasshopper.
@@
 These options however were not necessary for the level of optimization required within this project.
+Galapagos accepts Grasshopper slider components as input, with the domain set independently for each slider.
+As each DADU is reduced to a shoebox model, each facade is governed by an individual WWR slider, all with domain: $\lbrace x \vert 0 \leq x \leq 0.95 \rbrace$.
 
-%Together, the contextual 3D model, the specific design iteration of the DADU with varying WWRs, and the EUI value 
+Each Galapagos optimization was run twice for each site as a means to double check that the first result was not simply a local minimum. The evolutionary solver input panel was set as follows:
+
+% Itemized list of zoning restrictions for each site
+\begin{itemize}
+	\item Max. Stagnant = 50
+	\item Population = 75
+	\item Initial Boost = 4x
+	\item Maintain = 5\%
+	\item Inbreeding = +75\%
+\label{galapagos-settings}
+\end{itemize}
+
+Population set to 75 was found to be an effective balance between speed and reliability in initial tests.
+Allocating a 4x multiplier to initial population size ensures enough design variation from the start.
+Each optimization run was stopped at 15 generations, unless stagnation was reached earlier in the process.
+
 \section{Results}
-
-\section{Conclusion}
-%TODO Add main conclusion section
-%TODO Benefit of a pure python tool 
-Looking forward, moving on from the Grasshopper platform gives the benefit of non-reliance on developers to maintain the software in which the tool depends.
-Creation of such a tool in a singular programming language (in this case Python) further offers the ability to quickly and easily run the tool on a high variance of devices.
-In turn, this theoretically increases the rate of adoption by lowering the requirement to use the design tool.
-Additionally, Python is used within Rhino/Grasshopper (or a flavor thereof), provides many useful math and geometry libraries, and has options for injection into a web app (Flask and Django).
+%TODO introduce sites/addresses, site plans
+%TODO Add data visualizations of results
+%TODO additional simplified plans or renderings of DADUs and sites
+%TODO eui output breakdown and total, point out how only cooling and heating truly vary
 %TODO Limitations of Galapagos/Grasshopper workflow
-%TODO ***User ability to override window sizing and placement- feedback loop?
 Utilizing Grasshopper and Galapagos offers a ‘sandbox’ environment to begin to understand genetic algorithms and to arrive at tangible design solutions faster than using a home-brewed algorithm.
 Galapagos offers many fewer input parameters and a much narrower scope in which to define the fitness function.
-In its out-of-the-box form, Galapagos only accepts number sliders as input and can only optimize integers and floats- in reality, there is not means in which to define a true fitness function, only numerical values to target.
-%TODO future work of training machine learning model using energyplus data to replace simulation timesink
+In its out-of-the-box form, Galapagos only accepts number sliders as input and can only optimize integers and floats
+\section{Conclusion}
+%TODO Add main conclusion section, user ability to override window sizing and placement- feedback loop?
+Looking forward, the intention is to replace the computationally-expensive EnergyPlus simulation with a machine learning model that can predict EUI performance.
+This technique is essentially  surrogate modeling; that is, a means to approximate/predict the energy performance without doing the thermodynamics calculations required to return a true value.
+Machine learning algorithms require vast troves of data to learn from in order to increase accuracy past an acceptable threshold.
+This tool generates the optimized energy simulation data, which in turn becomes the training dataset.
+Assuming the resultant tool could be hosted on the web, this theoretically increases the rate of adoption by lowering the requirement to utilize the tool.
 
 %Figures are called in as follows:
-%\begin{verbatim}
 %\begin{figure}
@@
 %\caption{<Figure caption>}
+%\label{label}
 %\end{figure}
-%\end{verbatim}
-
-%\subsection{Cross-referencing}
-%The use of the \LaTeX\ cross-reference system
-%for figures, tables, equations, etc., is encouraged
-%(using \verb"\ref{<name>}" and \verb"\label{<name>}").
-%
-%\subsection{Endnotes}
-%Most \textit{SAGE} journals use endnotes rather than footnotes, so any notes should be coded as \verb+\endnote{<Text>}+.
-%Place the command \verb+\theendnotes+ just above the Reference section to typeset the endnotes.
-%
-%To avoid any confusion for papers that use Vancouver style references,  footnotes/endnotes should be edited into the text.
 
 \begin{acks}
-The author would like to thank Professor Tomas Mendez Echenagucia for his genetic algorithm expertise and input; Teresa Moroseos for her assistance with Honeybee debugging; Professor Alex Anderson and fellow students in Arch597 at University of Washington for feedback and motivation.
+The author would like to thank Professor Tomas Mendez Echenagucia for his genetic algorithm expertise and input; Teresa Moroseos for her assistance with Honeybee debugging; Professor Alex Anderson and fellow students in the Spring 2021 Arch597 - Research Practicum course at University of Washington for feedback and motivation; last but not least, Mia Lorentsen for her unending support but not for her assistance in procrastinating.
 \end{acks}

### 2021-06-12 writing/essay_latex/Energy_simulation_and_analysis_ADU.tex

@@
 Seattle, Washington; Portland, Oregon; and Vancouver, British Columbia are exploring the use of accessory dwelling units (ADUs) as one method to combat surging housing prices.
-Use of ADUs is an effective means of increasing housing density without replacing single family housing zones with new multi family residential construction.
+Use of ADUs is an effective means of increasing housing density without replacing single family (SF) housing zones with new multi family residential construction.
 Additionally, ADUs are often designed to be rented, generating supplementary income for the homeowner.
@@
 The ability to show users the estimated annual energy consumption of a new construction DADU would have a major impact on driving construction demand based on survey responses noted.
-Using an optimization method such as a genetic algorithm can allow architects to arrive at the most effective design solution on a site specific basis while predictably minimizing both environmental and financial costs.
+Using an optimization method such as a genetic algorithm (GA) can allow architects to arrive at the most effective design solution on a site specific basis while predictably minimizing both environmental and financial costs.
 
@@
 This climate type is utilized within the Grasshopper script to define the wood frame construction type for the Seattle climate.
+EPW files are obtainable freely at \textit{EnergyPlus.net}, the United States Department of Energy's official EnergyPlus website.
+
+Output data consists of the final, optimized site EUI value for each DADU as a sum value of cooling, heating, lighting, equipment, and hot water EUIs.
+Lighting, equipment, and hot water are all held constant at 1.923 kBTU/ft$^2$, 9.645 kBTU/ft$^2$, and 14.295 kBTU/ft$^2$, respectively.
 
 \subsection{Genetic algorithm}
-Genetic algorithms (GA) operate using logic that mimics biological evolution to optimize a value.
+Genetic algorithms operate using logic that mimics biological evolution to optimize a value.
 Fitness functions are the means by which to objectively evaluate the strength of individual designs.
@@
 Work was initially done using a purely Python-based GA script, but later pivoted to utilizing Galapagos, a genetic solver plugin that ships with Grasshopper.
-Galapagos offers a simple user interface and more direct integration within Grasshopper at the cost of advanced options.
+Utilizing Grasshopper and Galapagos offers a ‘sandbox’ environment to begin to understand genetic algorithms and to arrive at tangible design solutions faster than using a home-brewed algorithm.
+Further, Galapagos offers a simple user interface and a more direct integration within Grasshopper at the cost of advanced options.
 These options however were not necessary for the level of optimization required within this project.
 Galapagos accepts Grasshopper slider components as input, with the domain set independently for each slider.
-As each DADU is reduced to a shoebox model, each facade is governed by an individual WWR slider, all with domain: $\lbrace x \vert 0 \leq x \leq 0.95 \rbrace$.
+As each DADU is reduced to a shoebox model, each facade is governed by one individual WWR slider, each with domain $\lbrace x \vert 0 \leq x \leq 0.95 \rbrace$.
 
@@
 Allocating a 4x multiplier to initial population size ensures enough design variation from the start.
-Each optimization run was stopped at 15 generations, unless stagnation was reached earlier in the process.
+Each optimization run was stopped at 15 generations, with each site having been optimized 5 times.
+The results presented below for sites are averaged between all completed runs.
+Outlier EUI values were omitted and optimizations re-run.
+Ocassionally, Galapagos may reach what is termed a \textit{local minimum}, values which define the lowest EUI within the genetic dataset, but not the true minimal EUI attainable within constraints.
 
 \section{Results}
-%TODO introduce sites/addresses, site plans
+\begin{figure}[h!]
+\centering
+\includegraphics{site1.pdf}
+\caption{Site 1: 961 20th Ave, Seattle WA | DADU floor area: 1000 ft$^2$}
+\label{site1-plan}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics{site2.pdf}
+\caption{Site 2: 328 17th Ave, Seattle WA | DADU floor area: 520 ft$^2$}
+\label{site2-plan}
+\end{figure}
+
+Site selection was primarily driven by orientation between DADU location and existing residences, with each site in a differing neighborhood of Seattle.
+The Central District, Minor, and Madrona neighborhoods all lie east of downtown Seattle, greatly uphill.
+Neighborhoods differ slightly in density, street width, and demographics but all chosen sites are within identical SF5000 zoning.
+Site 1 is located at 961 20th Avenue in the Central District and the DADU is bordered directly by existing structures to the north and east.
+This DADU is the largest explored, reaching the maximum allowable floor area by single family code\cite{ADUniverseADURules}.
+Site 2 resides at 328 17th Avenue in Minor and its DADU structure is adjacent to the east of the existing home.
+This site is unique out of the three, as there is an additional existing structure located on the lot, being a small storage shed to the south-east.
+Finally, site 3, located at 3413 East Pine Street within Madrona consists of the existing residence facing the street to the north, with DADU location to the south.
+Neighboring residences to the east and southeast account for the majority of shading effect.
+
+\begin{figure}[h!]
+\centering
+\includegraphics{site3.pdf}
+\caption{Site 3: 3413 E Pine St, Seattle WA | DADU floor area: 680 ft$^2$}
+\label{site3-plan}
+\end{figure}
+
 %TODO Add data visualizations of results
@@
 %TODO Limitations of Galapagos/Grasshopper workflow
-Utilizing Grasshopper and Galapagos offers a ‘sandbox’ environment to begin to understand genetic algorithms and to arrive at tangible design solutions faster than using a home-brewed algorithm.
-Galapagos offers many fewer input parameters and a much narrower scope in which to define the fitness function.
-In its out-of-the-box form, Galapagos only accepts number sliders as input and can only optimize integers and floats
+
 \section{Conclusion}
@@
 
-%Figures are called in as follows:
-%\begin{figure}
-%\centering
-%\includegraphics{<figure name>}
-%\caption{<Figure caption>}
-%\label{label}
-%\end{figure}
-
 \begin{acks}
-The author would like to thank Professor Tomas Mendez Echenagucia for his genetic algorithm expertise and input; Teresa Moroseos for her assistance with Honeybee debugging; Professor Alex Anderson and fellow students in the Spring 2021 Arch597 - Research Practicum course at University of Washington for feedback and motivation; last but not least, Mia Lorentsen for her unending support but not for her assistance in procrastinating.
+The author would like to thank Professor Tomas Mendez Echenagucia for his genetic algorithm expertise and input; Teresa Moroseos for her assistance with Honeybee debugging; Professor Alex Anderson and fellow students in the Spring 2021 Arch597 - Research Practicum course at University of Washington for feedback and motivation; Mia Lorentsen for her unending support.
 \end{acks}

### 2021-06-14 writing/essay_latex/Energy_simulation_and_analysis_ADU.tex

@@
 
-\documentclass[sagev,times,Royal]{sagej}
+\documentclass[sagev,times,Review,doublespace]{sagej}
 % add 'doublespace'
 
-\usepackage{moreverb,url}
-
-\usepackage{graphicx}
+\usepackage{moreverb,url,graphicx}
 
@@
 \section{Introduction}
-The United States is amidst an unprecedented housing crisis, stemming from parallel crises: from neoliberal austerity spending cuts towards social benefits to the near-virtual lack of social housing and rent control programs in the United States to the ever-increasing privatization of essentially every aspect of American life\cite{10.2307/26297969}.
+The United States is amidst an unprecedented housing crisis, stemming from parallel crises: from neoliberal austerity spending cuts towards social benefits to the near-virtual lack of social housing and rent control programs in the US to the ever-increasing privatization of essentially every aspect of American life\cite{10.2307/26297969}.
 The result is a rising population of unhoused peoples and the inability for many to become a homeowner or to even afford minimum monthly rent in many cities.
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=4in]{3d-view.png}
+\caption{Isometric view of DADU, demonstrating WWR representation}
+\label{3d-view}
+\end{figure}
+
 Site locations were chosen through the use of the ADUniverse tool, developed by University of Washington researchers at the eScience Institute's \textit{Data Science for Social Good} program\cite{ADUniverseToolEScience}.
@@
 Under current Seattle Residential Building Code (2018), minimum R-values for wall and ceiling assemblies are R-21 and R-49, respectively.
-Light wood frame construction was chosen for all three sites, as the majority of residential construction in the United States and in Seattle specifically falls under this category due to price and ubiquity.
+Light wood frame construction was chosen for all three sites, as the majority of residential construction in the US and in Seattle specifically falls under this category due to price and ubiquity.
 For this project, total R-values were set to 26.86, 54.592, and 5.73, for wall, roof, and floor assemblies, respectively.
@@
 % Wall Assembly
-\begin{table}[]
+\begin{table}[h!]
 \centering
 \resizebox{\textwidth}{!}{%
-\begin{tabular}{lcccc}
-\hline
-\multicolumn{1}{|l|}{\textbf{Assembly}} &
-  \multicolumn{1}{l|}{\textbf{Notes}} &
-  \multicolumn{1}{c|}{\textbf{R-Value}} &
-  \multicolumn{1}{l|}{\textbf{Path 1 (Insulation)}} &
-  \multicolumn{1}{l|}{\textbf{Path 2 (Framing)}} \\ \hline
+\begin{tabular}{@{}lcccc@{}}
+\toprule
+\textbf{Assembly} &
+  \multicolumn{1}{l}{\textbf{Notes}} &
+  \textbf{R (\(\frac{^{\circ}F\,ft^2\,hr}{BTU}\))} &
+  \multicolumn{1}{l}{\textbf{Path 1} (Insulation)} &
+  \multicolumn{1}{l}{\textbf{Path 2} (Framing)} \\ \midrule
 \multicolumn{1}{|l|}{Air Film} &
@@
   \multicolumn{1}{c|}{0.68} &
-  \multicolumn{1}{c|}{0.68} \\ \hline
+  \multicolumn{1}{c|}{0.68} \\ \midrule
 \multicolumn{1}{|l|}{Gypsum Board} &
@@
   \multicolumn{1}{c|}{0.45} &
-  \multicolumn{1}{c|}{0.45} \\ \hline
+  \multicolumn{1}{c|}{0.45} \\ \midrule
 \multicolumn{1}{|l|}{Vapor Barrier} &
@@
   \multicolumn{1}{c|}{0} &
-  \multicolumn{1}{c|}{0} \\ \hline
+  \multicolumn{1}{c|}{0} \\ \midrule
 \multicolumn{1}{|l|}{Rigid Insulation} &
@@
   \multicolumn{1}{c|}{24} &
-  \multicolumn{1}{c|}{-} \\ \hline
+  \multicolumn{1}{c|}{-} \\ \midrule
 \multicolumn{1}{|l|}{2" x 6" Wood Stud} &
@@
   \multicolumn{1}{c|}{-} &
-  \multicolumn{1}{c|}{12.31} \\ \hline
+  \multicolumn{1}{c|}{12.31} \\ \midrule
 \multicolumn{1}{|l|}{Air Gap} &
@@
   \multicolumn{1}{c|}{2.38} &
-  \multicolumn{1}{c|}{2.38} \\ \hline
+  \multicolumn{1}{c|}{2.38} \\ \midrule
 \multicolumn{1}{|l|}{Wood Cladding} &
@@
   \multicolumn{1}{c|}{0.93} &
-  \multicolumn{1}{c|}{0.93} \\ \hline
+  \multicolumn{1}{c|}{0.93} \\ \midrule
 \multicolumn{1}{|l|}{Air Film} &
@@
   \multicolumn{1}{c|}{0.17} &
-  \multicolumn{1}{c|}{0.17} \\ \hline
+  \multicolumn{1}{c|}{0.17} \\ \midrule
  &
   \multicolumn{1}{c|}{} &
-  \multicolumn{1}{c|}{\textbf{R-Total:}} &
+  \multicolumn{1}{c|}{\textbf{R$_{total}$}} &
   \multicolumn{1}{c|}{28.61} &
-  \multicolumn{1}{c|}{16.92} \\ \cline{3-5} 
+  \multicolumn{1}{c|}{16.92} \\ \cmidrule(l){3-5} 
  &
@@
   24.3185 &
-  2.538 \\ \cline{3-5} 
+  2.538 \\ \cmidrule(l){3-5} 
  &
   \multicolumn{1}{l|}{} &
-  \multicolumn{1}{c|}{\textbf{Wall R-Value:}} &
-  \multicolumn{2}{c|}{26.86} \\ \cline{3-5} 
+  \multicolumn{1}{c|}{\textbf{R$_{wall}$}} &
+  \multicolumn{2}{c|}{26.86} \\ \cmidrule(l){3-5} 
  &
   \multicolumn{1}{l|}{} &
-  \multicolumn{1}{l|}{\textbf{Wall U-Value:}} &
-  \multicolumn{2}{c|}{0.0372} \\ \cline{3-5} 
+  \multicolumn{1}{c|}{\textbf{U$_{wall}$}} &
+  \multicolumn{2}{c|}{0.0372} \\
  &
@@
   \multicolumn{1}{l}{} &
-  \multicolumn{1}{l}{}
+  \multicolumn{1}{l}{} \\ \bottomrule
 \end{tabular}%
@@
 % Roof Assembly
-\begin{table}[]
+\begin{table}[h!]
 \centering
 \resizebox{\textwidth}{!}{%
-\begin{tabular}{lcccc}
-\hline
-\multicolumn{1}{|l|}{\textbf{Assembly}} &
-  \multicolumn{1}{l|}{\textbf{Notes}} &
-  \multicolumn{1}{c|}{\textbf{R-Value}} &
-  \multicolumn{1}{l|}{\textbf{Path 1 (Insulation)}} &
-  \multicolumn{1}{l|}{\textbf{Path 2 (Framing)}} \\ \hline
+\begin{tabular}{@{}lcccc@{}}
+\toprule
+\textbf{Assembly} &
+  \multicolumn{1}{l}{\textbf{Notes}} &
+  \textbf{R (\(\frac{^{\circ}F\,ft^2\,hr}{BTU}\))} &
+  \multicolumn{1}{l}{\textbf{Path 1} (Insulation)} &
+  \multicolumn{1}{l}{\textbf{Path 2} (Framing)} \\ \midrule
 \multicolumn{1}{|l|}{Air Film} &
@@
   \multicolumn{1}{c|}{0.61} &
-  \multicolumn{1}{c|}{0.61} \\ \hline
+  \multicolumn{1}{c|}{0.61} \\ \midrule
 \multicolumn{1}{|l|}{Gypsum Board} &
@@
   \multicolumn{1}{c|}{0} &
-  \multicolumn{1}{c|}{0} \\ \hline
-\multicolumn{1}{|l|}{Cellulose Insulation} &
-  \multicolumn{1}{c|}{Blow-in} &
-  \multicolumn{1}{c|}{47} &
-  \multicolumn{1}{c|}{47} &
-  \multicolumn{1}{c|}{-} \\ \hline
-\multicolumn{1}{|l|}{2" x 10" Wood Rafters} &
-  \multicolumn{1}{c|}{24" o.c.} &
-  \multicolumn{1}{c|}{14.68} &
-  \multicolumn{1}{c|}{-} &
-  \multicolumn{1}{c|}{14.68} \\ \hline
+  \multicolumn{1}{c|}{0} \\ \midrule
+\multicolumn{1}{|l|}{Cellulose Insulation}  & \multicolumn{1}{c|}{Blow-in}  & \multicolumn{1}{c|}{47}    & \multicolumn{1}{c|}{47}   & \multicolumn{1}{c|}{-}     \\ \midrule
+\multicolumn{1}{|l|}{2" x 10" Wood Rafters} & \multicolumn{1}{c|}{24" o.c.} & \multicolumn{1}{c|}{14.68} & \multicolumn{1}{c|}{-}    & \multicolumn{1}{c|}{14.68} \\ \midrule
 \multicolumn{1}{|l|}{Rigid Insulation} &
@@
   \multicolumn{1}{c|}{10} &
-  \multicolumn{1}{c|}{10} \\ \hline
-\multicolumn{1}{|l|}{OSB Sheathing} &
-  \multicolumn{1}{c|}{7/16"} &
-  \multicolumn{1}{c|}{0.77} &
-  \multicolumn{1}{c|}{0.77} &
-  \multicolumn{1}{c|}{0.77} \\ \hline
+  \multicolumn{1}{c|}{10} \\ \midrule
+\multicolumn{1}{|l|}{OSB Sheathing}         & \multicolumn{1}{c|}{7/16"}    & \multicolumn{1}{c|}{0.77}  & \multicolumn{1}{c|}{0.77} & \multicolumn{1}{c|}{0.77}  \\ \midrule
 \multicolumn{1}{|l|}{Asphalt Shingles} &
@@
   \multicolumn{1}{c|}{0.44} &
-  \multicolumn{1}{c|}{0.44} \\ \hline
+  \multicolumn{1}{c|}{0.44} \\ \midrule
 \multicolumn{1}{|l|}{Air Film} &
@@
   \multicolumn{1}{c|}{0.17} &
-  \multicolumn{1}{c|}{0.17} \\ \hline
+  \multicolumn{1}{c|}{0.17} \\ \midrule
  &
   \multicolumn{1}{c|}{} &
-  \multicolumn{1}{c|}{\textbf{R-Total:}} &
+  \multicolumn{1}{c|}{\textbf{R$_{total}$}} &
   \multicolumn{1}{c|}{59.44} &
-  \multicolumn{1}{c|}{27.12} \\ \cline{3-5} 
+  \multicolumn{1}{c|}{27.12} \\ \cmidrule(l){3-5} 
  &
@@
   50.524 &
-  4.068 \\ \cline{3-5} 
+  4.068 \\ \cmidrule(l){3-5} 
  &
   \multicolumn{1}{l|}{} &
-  \multicolumn{1}{c|}{\textbf{Wall R-Value:}} &
-  \multicolumn{2}{c|}{54.592} \\ \cline{3-5} 
+  \multicolumn{1}{c|}{\textbf{R$_{wall}$}} &
+  \multicolumn{2}{c|}{54.592} \\ \cmidrule(l){3-5} 
  &
   \multicolumn{1}{l|}{} &
-  \multicolumn{1}{l|}{\textbf{Wall U-Value:}} &
-  \multicolumn{2}{c|}{0.0183} \\ \cline{3-5} 
+  \multicolumn{1}{c|}{\textbf{U$_{wall}$}} &
+  \multicolumn{2}{c|}{0.0183} \\
  &
@@
 Honeybee, part of the open-source Ladybug Tools plugin suite for Grasshopper and Rhino, has become an industry standard for daylighting simulation and energy modeling within architecture and engineering, largely due to its open source nature\cite{mackey2018tool}.
-Honeybee acts as a bridge between modeled Rhino geometry and the United States Department of Energy's EnergyPlus simulation tool.
+Honeybee acts as a bridge between modeled Rhino geometry and the US Department of Energy's EnergyPlus simulation tool.
 Modeled geometry is constructed as rooms and zones within Honeybee, offering greater granularity.
@@
 Further, for Seattle specifically, results varied by only 2.12\%.
-
 Similarly, trees are represented as simple cylindrical solids, in order to keep energy simulation speed quick and to reduce the time necessary to reach an optimal solution.
@@
 This climate type is utilized within the Grasshopper script to define the wood frame construction type for the Seattle climate.
-EPW files are obtainable freely at \textit{EnergyPlus.net}, the United States Department of Energy's official EnergyPlus website.
-
+EPW files are obtainable freely at \textit{EnergyPlus.net}, the US Department of Energy's official EnergyPlus website.
 Output data consists of the final, optimized site EUI value for each DADU as a sum value of cooling, heating, lighting, equipment, and hot water EUIs.
@@
 Further, Galapagos offers a simple user interface and a more direct integration within Grasshopper at the cost of advanced options.
-These options however were not necessary for the level of optimization required within this project.
+However, these options were not necessary for the level of optimization required within this project.
 Galapagos accepts Grasshopper slider components as input, with the domain set independently for each slider.
 As each DADU is reduced to a shoebox model, each facade is governed by one individual WWR slider, each with domain $\lbrace x \vert 0 \leq x \leq 0.95 \rbrace$.
-
-Each Galapagos optimization was run twice for each site as a means to double check that the first result was not simply a local minimum. The evolutionary solver input panel was set as follows:
+Each Galapagos optimization was run five times for each site as a means to  check for outliers and not simply a local minimum.
+The evolutionary solver input panel was set as follows:
 
@@
 
-\section{Results}
 \begin{figure}[h!]
@@
 \includegraphics{site1.pdf}
-\caption{Site 1: 961 20th Ave, Seattle WA | DADU floor area: 1000 ft$^2$}
+\caption{Site 1 (Central District): 961 20th Avenue | DADU floor area: 1000 ft$^2$}
 \label{site1-plan}
 \end{figure}
+
+\section{Results}
+Site selection was primarily driven by orientation between DADU location and existing residences, with each site in a differing neighborhood of Seattle.
+The Central District, Minor, and Madrona neighborhoods all lie east of downtown Seattle, greatly uphill.
+Neighborhoods differ slightly in density, street width, and demographics but all chosen sites are within identical SF5000 zoning.
+Site 1 is located at 961 20th Avenue in the Central District and the DADU is bordered directly by existing structures to the north and east.
+This DADU is the largest explored, reaching the maximum allowable floor area of 1000 ft$^2$, as limited by single family code\cite{ADUniverseADURules}.
+Site 2 resides at 328 17th Avenue in Minor and its DADU structure is adjacent to the east of the existing home.
+This site is unique out of the three, as there is an additional existing structure located on the lot, being a small storage shed to the south-east.
+DADU footprint for site 2 totals 520 ft$^2$.
+Finally, site 3, located at 3413 East Pine Street within Madrona consists of the existing residence facing the street to the north, with DADU location to the south.
+Gross floor area of this DADU structure is 680 ft$^2$.
+Neighboring residences to the east and southeast account for the majority of shading effect.
+One future consideration is to allow for override to WWR domain, either automatically or manually selected by user.
+For example, let's say the DADU at site 3 found a value of 35\% for the southern facade.
+This may be optimal, but the majority of owners likely would opt to avoid windows facing directly into their neighbors private spaces of their dwelling.
+The genetic algorithm could contain a few toggles for override.
 
@@
 \includegraphics{site2.pdf}
-\caption{Site 2: 328 17th Ave, Seattle WA | DADU floor area: 520 ft$^2$}
+\caption{Site 2 (Minor): 328 17th Avenue | DADU floor area: 520 ft$^2$}
 \label{site2-plan}
@@
 
-Site selection was primarily driven by orientation between DADU location and existing residences, with each site in a differing neighborhood of Seattle.
-The Central District, Minor, and Madrona neighborhoods all lie east of downtown Seattle, greatly uphill.
-Neighborhoods differ slightly in density, street width, and demographics but all chosen sites are within identical SF5000 zoning.
-Site 1 is located at 961 20th Avenue in the Central District and the DADU is bordered directly by existing structures to the north and east.
-This DADU is the largest explored, reaching the maximum allowable floor area by single family code\cite{ADUniverseADURules}.
-Site 2 resides at 328 17th Avenue in Minor and its DADU structure is adjacent to the east of the existing home.
-This site is unique out of the three, as there is an additional existing structure located on the lot, being a small storage shed to the south-east.
-Finally, site 3, located at 3413 East Pine Street within Madrona consists of the existing residence facing the street to the north, with DADU location to the south.
-Neighboring residences to the east and southeast account for the majority of shading effect.
+Averaged optimal total site EUI was found to differ by less than 4 kBTU/ft$^2$ between all three sites.
+Total EUI was calculated to be 62.675 kBTU/ft$^2$, 61.207 kBTU/ft$^2$, and 64.059 kBTU/ft$^2$ for sites 1-3, respectively.
+Heating, on average, constituted for about 50\% of total EUI, while cooling accounted for only about 9\% of total site EUI on average.
+Optimal WWR values were found to be greatly similar between sites, regardless of context and orientation to existing residences.
+North, west, and east WWR values were typically found to lower EUI when lowered almost to zero.
+On the other hand, the south WWR approached .25 on average.
+North WWR was found to fall between 0 and 0.18, with an average value of 0.08 percent glazed to opaque.
+East and west WWRs typically fell between 0 and 0.11, with the majority of Galapagos runs ending with zero glazing on the east and west facades.
+Curiously, the lowest EUI found out of any runs was located at site 2 in Minor, with WWRs of 0.06 north, 0.09 east, 0.10 south, and 0 west.
+This combination of window to wall ratios stands out as it was simultaneously the only resultant where north, east, and south ratios are almost identical and non-zero.
 
@@
 \includegraphics{site3.pdf}
-\caption{Site 3: 3413 E Pine St, Seattle WA | DADU floor area: 680 ft$^2$}
+\caption{Site 3 (Madrona): 3413 East Pine Street | DADU floor area: 680 ft$^2$}
 \label{site3-plan}
@@
 
-%TODO Add data visualizations of results
-%TODO additional simplified plans or renderings of DADUs and sites
-%TODO eui output breakdown and total, point out how only cooling and heating truly vary
-%TODO Limitations of Galapagos/Grasshopper workflow
+Table 3 displays the averaged results by site, as well as overall.
+Comparing these optimal values to a worst case and an educated guess demonstrates the value in using such a relatively simple design tool like Galapagos- one which is already bundled with Rhinoceros software (as of version 6).
+A worst case result was calculated for site 3 by maximizing the EUI using Galapagos.
+Total site EUI was found to be 91.048 kBTU/ft$^2$, a 42.13\% increase over the optimal solution's total site EUI value of 64.059 kBTU/ft$^2$.
+North, east, south, and west WWRs were set by the solver to 0.79, 0.95, 0.95, and 0.90, respectively.
+An educated guess consisting of WWR settings of 0.06, 0.18, 0.59, and 0.18, in the aforementioned order leads to a total site EUI of 69.643 kBTU/ft$^2$.
+In comparison to the optimal WWR arrangement, this results in only a 8.717\% EUI increase.
+This result can be taken to mean that even for seasoned designers, this tool can still offer nearly 10\% greater energy efficiency, as well as a simple way to verify hand calculations or other means of estimating.
+
+\begin{table}[h!]
+\centering
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{@{}lcccccc@{}}
+\toprule
+ &
+  \multicolumn{6}{c}{\textbf{Energy Use Intensity (kBTU/ft$^2$)}} \\ \midrule
+\multicolumn{1}{|c|}{Site} &
+  \multicolumn{1}{c|}{Cooling} &
+  \multicolumn{1}{c|}{Heating} &
+  \multicolumn{1}{c|}{Lighting} &
+  \multicolumn{1}{c|}{Equipment} &
+  \multicolumn{1}{c|}{Hot Water} &
+  \multicolumn{1}{c|}{Total} \\ \midrule
+\multicolumn{1}{|l|}{961 20th Ave} &
+  \multicolumn{1}{c|}{4.169} &
+  \multicolumn{1}{c|}{32.643} &
+  \multicolumn{1}{c|}{1.923} &
+  \multicolumn{1}{c|}{9.645} &
+  \multicolumn{1}{c|}{14.295} &
+  \multicolumn{1}{c|}{62.675} \\ \midrule
+\multicolumn{1}{|l|}{328 17th Ave} &
+  \multicolumn{1}{c|}{7.657} &
+  \multicolumn{1}{c|}{27.688} &
+  \multicolumn{1}{c|}{1.923} &
+  \multicolumn{1}{c|}{9.645} &
+  \multicolumn{1}{c|}{14.295} &
+  \multicolumn{1}{c|}{61.207} \\ \midrule
+\multicolumn{1}{|l|}{3413 E Pine St} &
+  \multicolumn{1}{c|}{5.039} &
+  \multicolumn{1}{c|}{33.157} &
+  \multicolumn{1}{c|}{1.923} &
+  \multicolumn{1}{c|}{9.645} &
+  \multicolumn{1}{c|}{14.295} &
+  \multicolumn{1}{c|}{64.059} \\ \midrule
+ &
+  \multicolumn{1}{l}{} &
+  \multicolumn{1}{l}{} &
+  \multicolumn{1}{l}{} &
+  \multicolumn{1}{l}{} &
+  \multicolumn{1}{l}{} &
+  \multicolumn{1}{l}{} \\
+\textbf{Average} &
+  5.621 &
+  31.162 &
+  1.923 &
+  9.645 &
+  14.295 &
+  62.647 \\ \bottomrule
+\end{tabular}%
+}
+\label{results-table}
+\caption{Optimized EUI Results}
+\end{table}
 
 \section{Conclusion}
-%TODO Add main conclusion section, user ability to override window sizing and placement- feedback loop?
+ADUs are an effective way to curtail housing rental costs by increasing housing density without fundamentally changing the housing landscape.
+ADUs further offer homeowners a means of passive income or compelling alternative to nursing homes for elderly parents and other relatives.
+Reducing energy use intensity for accessory dwelling unit structures early in the design process offers the prospect to assist in increasing ADU uptake and construction rate.
+By automating the process using a genetic search algorithm and an energy simulation tool in tandem, experienced architects and designers can reduce total site EUI by an average of 8.018\%.
+Inexperienced designers, or those without ballpark window to wall ratios to work with could see up to (theoretically) a 29.643\% reduction in EUI based on experimental results.
+Regardless, such a tool can be utilized to verify EUI estimates during the schematic design stage.
+
 Looking forward, the intention is to replace the computationally-expensive EnergyPlus simulation with a machine learning model that can predict EUI performance.

### 2022-02-22 writing/ACADIA21/essay_latex/Energy_simulation_and_analysis_ADU.tex
paper edits

--- 2021-06-14 writing/essay_latex/Energy_simulation_and_analysis_ADU.tex
@@
 The effect of varying window-to-wall ratios (WWRs) on EUI is simulated using EnergyPlus and Honeybee.
-Galapagos, a genetic solver plugin for Grasshopper is used to vary window-to-wall ratio and optimize ADU energy use intensity on three single family residential lots in Seattle.
+Galapagos, a genetic solver plugin for Grasshopper, is used to vary window-to-wall ratio and optimize ADU energy use intensity on three single family residential lots in Seattle.
 Site context is constructed using OpenStreetMap GIS data and zoning restrictions are incorporated into the Grasshopper definition.
@@
 \section{Introduction}
-The United States is amidst an unprecedented housing crisis, stemming from parallel crises: from neoliberal austerity spending cuts towards social benefits to the near-virtual lack of social housing and rent control programs in the US to the ever-increasing privatization of essentially every aspect of American life\cite{10.2307/26297969}.
-The result is a rising population of unhoused peoples and the inability for many to become a homeowner or to even afford minimum monthly rent in many cities.
+The United States is amidst an unprecedented housing crisis, stemming from parallel crises.
+From neoliberal austerity spending cuts towards social benefits to the near-virtual lack of social housing and rent control programs in the US to the ever-increasing privatization of essentially every aspect of American life\cite{10.2307/26297969}, housing is becoming more of a privilege.
+The result is a rising population of unhoused peoples and the inability for many to become a homeowner or to even afford minimum monthly rent in most top US cities.
 Solutions to this dilemma are neither straightforward nor definite.
@@
 This survey was positioned before the official call for submissions was broadcast.
-'Low cost' was the top ranked criteria with 48\% of responses listing 'very important', followed by 'green building' being ranked similarly at 46\%.
+Low cost was the top ranked criteria with 48\% of responses listing 'very important', followed by green building being ranked similarly at 46\%.
 Other suggestions sourced from surveyed respondents included low environmental costs, site specific considerations, and predictability.
@@
 Modeled geometry is constructed as rooms and zones within Honeybee, offering greater granularity.
-EnergyPlus operates using various abstractions and simplifications and these are important to note\cite{EnergyPlusTMVersionDocumentation2021}.
+Importantly, EnergyPlus operates using various abstractions and simplifications\cite{EnergyPlusTMVersionDocumentation2021}.
 Each zone and surface is only able to hold one temperature value at a time- each singular value is averaged across the zone or surface.

### 2022-02-22 writing/ACADIA21/essay_latex/chapters/conclusion.tex
paper edits

--- 2021-06-10 writing/essay_latex/draft3.tex
@@
-% sage_latex_guidelines.tex V1.20, 14 January 2017
+\chapter{Conclusion}
+\label{chap:conclusion}
 
-\documentclass[sagev,times,Royal]{sagej}
-% add 'doublespace'
+Benefit of a pure python tool
 
-\usepackage{moreverb,url}
-
-\usepackage{graphicx}
-
-\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}
-
-\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
-T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-
-\def\volumeyear{2021}
-
-\begin{document}
-
-\runninghead{Pape}
-
-\title{Energy simulation and analysis of accessory dwelling units: 
-an evolutionary approach}
-
-\author{P. Arthur Pape\affilnum{1}}
-
-\affiliation{\affilnum{1}University of Washington}
-
-\corrauth{P. Arthur Pape}
-
-\email{ppape@uw.edu}
-
-\begin{abstract}
-Accessory dwelling units (ADUs) are surging in popularity as cities push for their construction to aid in combatting the housing crisis.
-In Seattle, recent survey data calls for increased focus on sustainability and cost.
-Lowering energy use intensity (EUI) of ADUs allows for the targeting of both desires which homeowners hold.
-The effect of varying window-to-wall ratios (WWRs) on EUI is simulated using EnergyPlus and Honeybee.
-Galapagos, a genetic solver plugin for Grasshopper is used to vary window-to-wall ratio and optimize ADU energy use intensity on three single family residential lots in Seattle, Washington.
-Site context is constructed using OpenStreetMap GIS data and zoning restrictions are incorporated into the Grasshopper definition.
-Future work is planned to utilize output data from this tool as the training dataset for a machine learning model.
-Accurate machine learning prediction of energy use intensity instead of extended simulation would increase pragmatic use of such tools.
-\end{abstract}
-
-\keywords{Genetic algorithm, accessory dwelling unit, housing, optimization, simulation, energy use intensity, grasshopper, galapagos}
-
-\maketitle
-
-%TODO Complete explanation of Honeybee energy model and Galapagos genetic solver subsections within methods.
-%TODO Add data visualizations of results, additional simplified plans or renderings of DADUs and sites
-%TODO Flesh out conclusion, limit and refine portion on future work towards machine learning surrogate modeling
-
-\section{Introduction}
-The United States is amidst an unprecedented housing crisis, stemming from parallel crises: from neoliberal austerity spending cuts towards social benefits to the near-virtual lack of social housing and rent control programs in the United States to the ever-increasing privatization of essentially every aspect of American life\cite{10.2307/26297969}.
-The result is a rising population of unhoused peoples and the inability for many to become a homeowner or to even afford minimum monthly rent in many cities.
-Solutions to this dilemma are neither straightforward nor definite.
-However, designers and architects can only address the symptom-level problems, while advocating towards change directed at the root cause.
-
-Seattle, Washington; Portland, Oregon; and Vancouver, British Columbia are exploring the use of accessory dwelling units (ADUs) as one method to combat surging housing prices.
-Use of ADUs is an effective means of increasing housing density without replacing single family housing zones with new multi family residential construction.
-Additionally, ADUs are often designed to be rented, generating supplementary income for the homeowner.
-Historically, homeowner’s associations and other local organizations have  been resistant to any proposed density increases through zoning or other means\cite{10.2307/24392672}.
-However, Seattle and the other previously mentioned cities in the Pacific Northwest have succeeded in allowing for the construction of ADUs in recent decades, with these groups beginning to see ADUs as a positive.
-ADUs offer many positives beyond an increased housing supply such as a path for homeowners to generate income and accommodating households with multigenerational or caretaking needs.
-
-Beginning in 2014, the City of Seattle has moved to increase production of ADUs through various resolutions and legislation, in turn removing barriers to entry and highlighting strategies for encouragement.
-In late 2019, a call for detached ADU (or DADU) designs was issued to designers.
-As a result, ten DADU designs were pre-approved by the City of Seattle and are available online to entice homeowners\cite{ADUniversePreapprovedADU}.
-The pre-approved designs are free, but still require standard permitting fees to be paid for approval and come with one significant downside of many other pre-designed structures: a lack of contextual design and individualization.
-This one-size-fits-all approach can have drastic effects upon energy consumption, reducing the appeal of such an investment.
-Further, these designs do not offer the scalability or energy efficiency that an ADU designed specifically per site offers.
-
- Coincidentally, a November 2019 city survey shows that there are calls for an increased focus on sustainability and cost in the construction of ADUs\cite{seattlePreapprovedPlansAccessory2019}.
-This survey was positioned before the official call for submissions was broadcast.
-'Low cost' was the top ranked criteria with 48\% of responses listing 'very important', followed by 'green building' being ranked similarly at 46\%.
-Other suggestions sourced from surveyed respondents included low environmental costs, site specific considerations, and predictability.
-The ability to show users the estimated annual energy consumption of a new construction DADU would have a major impact on driving construction demand based on survey responses noted.
-Using an optimization method such as a genetic algorithm can allow architects to arrive at the most effective design solution on a site specific basis while predictably minimizing both environmental and financial costs.
-
-\section{Methods} 
-\subsection{Design constraints and decisions}
-This research explores the use of a genetic solver to find the optimal design solution of window-to-wall ratios (WWR) for a simple DADU form resulting in the lowest EUI value.
-Three sites in similar neighborhoods in the city of Seattle were selected based on a combination of proximity and slight density variance.
-The tool adheres to Seattle's land use code to find estimated buildable area for a DADU on each site, creates a simple box-like form representing the DADU and then varies the WWR value for each of the four walls.
-A genetic solver plugin is then allowed to vary these ratios to find an optimal solution using a process largely mimicking natural genetic processes.
-Finally, an energy model is created from each individual iteration to find an estimate for overall EUI.
-From here, the process loops until an optimal solution (in this case, via minimizing EUI) is found.
-
-%TODO merge these two (upper and lower) sections together
-Site locations were chosen through the use of the \textit{ADUniverse} tool, developed by the University of Washington researchers at the eScience Institute\cite{ADUniverseToolEScience}.
-ADUniverse allows for users to investigate the possibility of constructing an ADU through a map-based user interface.
-Chosen sites are located within the Madrona, Minor, and Central District neighborhoods of Seattle, WA.
-All three sites are zoned under SF5000, a single family residential zoning.
-Site data was extracted from the city of Seattle's GIS website, converted into an AutoCAD .DWG file, and then imported into and extruded within Rhinoceros 6 (Rhino), a 3D modeling software.
-Lot lines surrounding each house are fed into the Grasshopper definition, coupled with constraints in line with the current Seattle Land Use Code.
-Such universal constraints for ADUs on lots located within SF5000 zoning are as follows\cite{ADURules}:
-
-% itemized list of zoning restrictions for each site
-\begin{itemize}
-	\item Lot size must be at least 3200 square feet
-	\item Minimum lot width of 25 feet
-	\item Minimum lot length of 70 feet
-	\item DADU can occupy \textit{at most} 60\% of rear yard	
-	\item Front yard of 20 feet, side yards of 5 feet, rear yard of 25  feet
-	\item DADU height limit varies based on lot width: \begin{itemize}
-	\item 14 feet for lot widths less than 30 feet
-	\item 16 feet for lot widths between 40 and 50 feet
-	\item 18 feet for lot widths exceeding 50 feet \end{itemize}
-\end{itemize}
-
-These constraints are built into the Grasshopper definition in such a way that the tool only requires lot extents and the outline of the existing residential structure to develop the buildable area for any DADU construction.
-The DADU is represented as a simple box with dimensions of (fill this in once all 3 sites have been finalized).
-This minimal geometry is located congruent to lot lines and the existing structure.
-Under current Seattle Residential Building Code (2018), minimum R-values for wall and ceiling assemblies are R-21 and R-49, respectively.
-Light wood frame construction was chosen for all three sites, as the majority of residential construction in the United States and in Seattle specifically falls under this category due to price and ubiquity.
-For this project, wall assembly was specified to a total R-value of 29.232 (see Table \ref{table:wall-assembly}.).
-\textit{\textbf{***Insert second table or append data to existing table of slab and ceiling R-values***}}
-
-Next, the variable aspect of the DADU lies in the window-to-wall ratios of each of the four facades, which line up with the cardinal directions.
-The ratio between solid and glazed facade areas is allowed to shift between
-
-%TODO revise assembly table to match assembly used in grasshopper
-\begin{table}[]
-\centering
-\resizebox{\textwidth}{!}{%
-\begin{tabular}{lcccc}
-\hline
-\multicolumn{1}{|l|}{\textbf{Assembly}} &
-  \multicolumn{1}{l|}{\textbf{Notes}} &
-  \multicolumn{1}{c|}{\textbf{R-Value}} &
-  \multicolumn{1}{l|}{\textbf{Path 1 (Insulation)}} &
-  \multicolumn{1}{l|}{\textbf{Path 2 (Framing)}} \\ \hline
-\multicolumn{1}{|l|}{Air Film} &
-  \multicolumn{1}{c|}{Interior} &
-  \multicolumn{1}{c|}{0.68} &
-  \multicolumn{1}{c|}{0.68} &
-  \multicolumn{1}{c|}{0.68} \\ \hline
-\multicolumn{1}{|l|}{Gypsum Board} &
-  \multicolumn{1}{c|}{1/2"} &
-  \multicolumn{1}{c|}{0.45} &
-  \multicolumn{1}{c|}{0.45} &
-  \multicolumn{1}{c|}{0.45} \\ \hline
-\multicolumn{1}{|l|}{Vapor Barrier} &
-  \multicolumn{1}{c|}{} &
-  \multicolumn{1}{c|}{0} &
-  \multicolumn{1}{c|}{0} &
-  \multicolumn{1}{c|}{0} \\ \hline
-\multicolumn{1}{|l|}{Fiberglass Insulation} &
-  \multicolumn{1}{c|}{} &
-  \multicolumn{1}{c|}{3.5} &
-  \multicolumn{1}{c|}{28} &
-  \multicolumn{1}{c|}{-} \\ \hline
-\multicolumn{1}{|l|}{2" x 8" Wood Stud} &
-  \multicolumn{1}{c|}{16" o.c.} &
-  \multicolumn{1}{c|}{14.68} &
-  \multicolumn{1}{c|}{-} &
-  \multicolumn{1}{c|}{14.68} \\ \hline
-\multicolumn{1}{|l|}{Air Gap} &
-  \multicolumn{1}{c|}{1"} &
-  \multicolumn{1}{c|}{1} &
-  \multicolumn{1}{c|}{1} &
-  \multicolumn{1}{c|}{1} \\ \hline
-\multicolumn{1}{|l|}{Wood Cladding} &
-  \multicolumn{1}{c|}{3/4"} &
-  \multicolumn{1}{c|}{0.93} &
-  \multicolumn{1}{c|}{0.93} &
-  \multicolumn{1}{c|}{0.93} \\ \hline
-\multicolumn{1}{|l|}{Air Film} &
-  \multicolumn{1}{c|}{Exterior} &
-  \multicolumn{1}{c|}{0.17} &
-  \multicolumn{1}{c|}{0.17} &
-  \multicolumn{1}{c|}{0.17} \\ \hline
- &
-  \multicolumn{1}{c|}{} &
-  \multicolumn{1}{c|}{\textbf{R-Total:}} &
-  \multicolumn{1}{c|}{31.23} &
-  \multicolumn{1}{c|}{17.91} \\ \cline{3-5} 
- &
-  \multicolumn{1}{l}{} &
-  \multicolumn{1}{l}{} &
-  \textit{x 85\%} &
-  \textit{x 15\%} \\
- &
-  \multicolumn{1}{l}{} &
-  \multicolumn{1}{l}{} &
-  26.5455 &
-  2.6865 \\ \cline{3-5} 
- &
-  \multicolumn{1}{l|}{} &
-  \multicolumn{1}{c|}{\textbf{Wall R-Value:}} &
-  \multicolumn{2}{c|}{29.232} \\ \cline{3-5} 
- &
-  \multicolumn{1}{l|}{} &
-  \multicolumn{1}{l|}{\textbf{Wall U-Value:}} &
-  \multicolumn{2}{c|}{0.0342} \\ \cline{3-5} 
- &
-  \multicolumn{1}{l}{} &
-  \multicolumn{1}{l}{} &
-  \multicolumn{1}{l}{} &
-  \multicolumn{1}{l}{}
-\end{tabular}%
-}
-\caption{Wall Assembly R- and U-Values}
-\label{table:wall-assembly}
-\end{table}
-
-\subsection{Energy model and analysis}
-Honeybee, part of the open-source Ladybug Tools plugin suite for Grasshopper and Rhino, has become an industry standard for daylighting simulation and energy modeling within architecture and engineering.
-Honeybee acts as a bridge between modeled Rhino geometry and the United States Department of Energy's EnergyPlus simulation tool.
-Modeled geometry is constructed as rooms and zones within Honeybee, offering greater granularity.
-Using four zones and a core allows Honeybee and EnergyPlus to take into account the effect of glazing proportions on each facade has on heating and cooling EUI in more detail.
-EnergyPlus operates using various abstractions and simplifications and these are important to note.
-Each zone and surface is only able to hold one temperature value at a time- each singular value is averaged across the zone or surface.
-Each DADU simulated is represented as four zones and a core, the standard for simulations of this type.
-Alternatively, using one zone instead would have assisted with quicker simulation times, and therefore quicker optimization speed.
-
-Similarly, trees are represented as simple cylindrical solids, in order to keep energy simulation speed quick and to reduce the time necessary to reach an optimal solution.
-In initial tests, trees were represented as spherical meshes created using Grasshopper.
-These more-detailed trees varied in opaque triangulated mesh coverage to simulate foliage cover and to more accurately cast shadows.
-It is important to note that with access to greater computing power it could become reasonable to represent and simulate trees in the original format.
-
-%TODO all inputs, geometry or otherwise required for honeybee/energyplus
-
-\subsection{Genetic algorithm}
-Genetic algorithms (GA) operate using logic that mimics biological evolution to optimize a value.
-Fitness functions are the means by which to objectively evaluate the strength of individual designs.
-In this case, the fitness function specified works to minimize the resultant EUI value of each DADU design.
-Work was initially done using a purely Python-based GA script, but later pivoted to utilizing Galapagos, a genetic solver plugin that ships with Grasshopper.
-Galapagos offers a simple user interface and more direct integration within Grasshopper at the cost of advanced options.
-These options however were not necessary for the level of optimization required within this project.
-
-%Together, the contextual 3D model, the specific design iteration of the DADU with varying WWRs, and the EUI value 
-\section{Results}
-
-\section{Conclusion}
-%TODO Add main conclusion section
-%TODO Benefit of a pure python tool 
 Looking forward, moving on from the Grasshopper platform gives the benefit of non-reliance on developers to maintain the software in which the tool depends.
@@
 Additionally, Python is used within Rhino/Grasshopper (or a flavor thereof), provides many useful math and geometry libraries, and has options for injection into a web app (Flask and Django).
-%TODO Limitations of Galapagos/Grasshopper workflow
-%TODO ***User ability to override window sizing and placement- feedback loop?
-Utilizing Grasshopper and Galapagos offers a ‘sandbox’ environment to begin to understand genetic algorithms and to arrive at tangible design solutions faster than using a home-brewed algorithm.
+
+Limitations of Galapagos/Grasshopper workflow
+
+Utilizing Galapagos and Grasshopper offers a ‘sandbox’ environment to begin to understand genetic algorithms and to arrive at tangible design solutions faster than using a home-brewed algorithm.
 Galapagos offers many fewer input parameters and a much narrower scope in which to define the fitness function.
 In its out-of-the-box form, Galapagos only accepts number sliders as input and can only optimize integers and floats- in reality, there is not means in which to define a true fitness function, only numerical values to target.
-%TODO future work of training machine learning model using energyplus data to replace simulation timesink
-
-%Figures are called in as follows:
-%\begin{verbatim}
-%\begin{figure}
-%\centering
-%\includegraphics{<figure name>}
-%\caption{<Figure caption>}
-%\end{figure}
-%\end{verbatim}
-
-%\subsection{Cross-referencing}
-%The use of the \LaTeX\ cross-reference system
-%for figures, tables, equations, etc., is encouraged
-%(using \verb"\ref{<name>}" and \verb"\label{<name>}").
-%
-%\subsection{Endnotes}
-%Most \textit{SAGE} journals use endnotes rather than footnotes, so any notes should be coded as \verb+\endnote{<Text>}+.
-%Place the command \verb+\theendnotes+ just above the Reference section to typeset the endnotes.
-%
-%To avoid any confusion for papers that use Vancouver style references,  footnotes/endnotes should be edited into the text.
-
-\begin{acks}
-The author would like to thank Professor Tomas Mendez Echenagucia for his genetic algorithm expertise and input; Teresa Moroseos for her assistance with Honeybee debugging; Professor Alex Anderson and fellow students in Arch597 at University of Washington for feedback and motivation.
-\end{acks}
-
-\begin{funding}
-The author(s) received no financial support for the research, authorship, and/or publication of this article.
-\end{funding}
-
-\begin{dci}
-	The author(s) declared no potential conflicts of interest with respect to the research, authorship, and/or publication of this article.
-\end{dci}
-
-\bibliographystyle{SageV}
-\bibliography{597_bibliography}
-
-\end{document}

### 2022-02-22 writing/ACADIA21/essay_latex/chapters/introduction.tex
paper edits

\chapter{Introduction}
\label{chap:intro}

The United States is amid an unprecedented housing crisis, stemming from issues outside of housing, such as neoliberal cuts to social benefits spending and the increasing privatization of essentially every aspect of American life. The result is a rising population of unhoused peoples and the inability for many to become a homeowner or to even afford minimum monthly rent in many cities. Solutions to this dilemma are neither straightforward nor definite. However, as designers and architects, this problem can only be addressed at the symptom-level, while advocating at the root cause.

%TODO Expand background portion of intro

Seattle, Washington; Portland, Oregon; and Vancouver, British Columbia are exploring the use of accessory dwelling units (ADUs) as one method to combat surging housing prices. Use of ADUs is an effective means of increasing housing density without replacing single family housing zones with new multi family residential construction. Additionally, ADUs are often designed to be rented, generating supplementary income for the homeowner. Ten detached ADU (or DADU) designs are pre-approved by the City of Seattle and are available online to entice homeowners. However, a 2019 city survey shows that there are calls for an increased focus on sustainability and cost in the construction of ADUs\cite{seattlePreapprovedPlansAccessory2019}. These out of the box designs do not offer the scalability or energy efficiency that an ADU designed specifically per site offers.

%TODO Expand reason for why this tool is necessary to drive ADU demand

Homeowner’s associations and other local organizations have historically fought back against any proposed density increases through zoning or other method. However, Seattle and the other aforementioned cities in the Pacific Northwest have succeeded in allowing for the construction of ADUs in recent decades. The pre-approved designs are free, but require payment of around one thousand dollars [check this] for approval and come with one significant downside of many other pre-designed structures- a lack of contextual design and individualization.

This research explores whether the use of genetic algorithms via shape grammar methodology can effectively optimize DADU plans to site context, increases building performance or further encourages construction. The proposed methodology begins by reading example site data from the city of Seattle including building and vegetation context from the Seattle GIS. Next, a genetic algorithm explores the design space for a viable floor plan solution based on a fitness function. This fitness function evaluates individual designs according to predefined traits. Traits to evaluate include window to wall ratio, insulation depth/type, ventilation strategy, and shading technique. Locating the correct combination of traits to minimize (or maximize) which results in a higher performance DADU is the desired outcome. Resulting designs will be analyzed and compared via energy performance simulation. The end-goal is to develop a computational tool using the aforementioned system to conduct automated site analysis, as well as parametric generation of DADUs.

### 2022-02-22 writing/ACADIA21/essay_latex/examples via tomas/chapter_TNA/TNA.tex
paper edits

\chapter{Thrust Network Analysis}
\label{chap:tna}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

Figure \ref{fig:texas_model} in page \pageref{fig:texas_model} is cool and i
have nothing else to say about it. Do you like it?

\begin{equation}
D_{50} =   \frac{\displaystyle{\int^{50}_{0}p^{2}(t)
\mathrm{d}t}}{\displaystyle{\int^{\infty}_{0}p^{2}(t) \mathrm{d}t}}   \quad \mathrm{(dB)}
\end{equation}

\section{Some Stuff}
\label{sec:some_stuff}

Deb \cite{Deb2001} talks about GA's and blah blah blah\ldots. He probably said
something like this:

\begin{quotation}
``Time and motion studies show that the fewer kinds of materials there are, the more efficient factory assembly is - and therefore demand a certain simplicity in the variety of materials used. This need for simplicity conflicts with the fact that the form will function better if we choose the best material for each separate purpose separately. But then, on the other hand, functional diversity of materials makes for expensive and complicated joints between components, which is liable to make maintenance less easy. Further still, all three issues, simplicity, performance and jointing, are at odds with our  to minimize the cost of materials\cite{tms2013c}. For if we choose the cheapest material for each separate task, we shall not necessarily have simplicity, nor optimum performance, nor materials which can be cleanly jointed.''
\end{quotation}

 Lorem ipsum dolor  sit amet, consectetur adipiscing elit, sed do
 eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

  Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
 incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\begin{figure}[ht]%[htbp]
\begin{center}
\begin{center}\includegraphics[width = 1\textwidth]{texas_model.jpg}\end{center}
\caption{TNA dual graph of the MLK memorial blah blah\ldots}
\label{fig:texas_model}
\end{center}
\end{figure}

  Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
 incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

  Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
 incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

  Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
 incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

 This is some pseudocode example from an awesome PhD thesis by Tom\'{a}s
 M\'{e}ndez Echenagucia:\\

 \begin{algorithmic}[1]
\State Generate a random and coded population 
\For{$i\gets 1$, number of Generations}
    \State Decode and Scale the population
    \State Calculate fitness values for the population
    \State Run Selection operation
    \State Run Crossover operation
    \State Run Mutation operation
    \State Perform Exit condition test

    \If{Exit condition test = \texttt{True}}
        \State \texttt{Exit} GA
    \ElsIf{Exit condition test = \texttt{False}}
        \State Continue to the next Generation
    \EndIf
\EndFor
\end{algorithmic}

I told you it was awesome.

### 2022-02-22 writing/ACADIA21/essay_latex/examples via tomas/chapter_tassellation/tassellation.tex
paper edits

--- 2022-02-22 writing/ACADIA21/essay_latex/examples via tomas/chapter_TNA/TNA.tex
@@
-\chapter{Thrust Network Analysis}
-\label{chap:tna}
+\chapter{Cool Tassellation}
+\label{chap:tass}
+
+\begin{figure}[ht]%[htbp]
+\begin{center}
+\begin{center}\includegraphics[width =
+1\textwidth]{texas_tassellation.jpg}\end{center}
+\caption{Model of the MLK memorial blah blah\ldots}
+\label{fig:texas_tassellation}
+\end{center}
+\end{figure}
 
@@
 
-Figure \ref{fig:texas_model} in page \pageref{fig:texas_model} is cool and i
-have nothing else to say about it. Do you like it?
+\begin{itemize}
+\item $P_{(0,0)}$ represents the completely flat surface, it has a maximum displacement in the z-axis (our fitness value) of 285 mm;
+\item $P_{(26,1)}$ has a maximum displacement in z of 21 mm; 
+\item $P_{(40, 40)}$ has a maximum displacement in z of 43 mm.
+\end{itemize}
 
-\begin{equation}
-D_{50} =   \frac{\displaystyle{\int^{50}_{0}p^{2}(t)
-\mathrm{d}t}}{\displaystyle{\int^{\infty}_{0}p^{2}(t) \mathrm{d}t}}   \quad \mathrm{(dB)}
-\end{equation}
-
-\section{Some Stuff}
-\label{sec:some_stuff}
-
-Deb \cite{Deb2001} talks about GA's and blah blah blah\ldots. He probably said
-something like this:
-
-\begin{quotation}
-``Time and motion studies show that the fewer kinds of materials there are, the more efficient factory assembly is - and therefore demand a certain simplicity in the variety of materials used.
-This need for simplicity conflicts with the fact that the form will function better if we choose the best material for each separate purpose separately.
-But then, on the other hand, functional diversity of materials makes for expensive and complicated joints between components, which is liable to make maintenance less easy.
-Further still, all three issues, simplicity, performance and jointing, are at odds with our  to minimize the cost of materials\cite{tms2013c}.
-For if we choose the cheapest material for each separate task, we shall not necessarily have simplicity, nor optimum performance, nor materials which can be cleanly jointed.''
-\end{quotation}
-
- Lorem ipsum dolor  sit amet, consectetur adipiscing elit, sed do
- eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\section{Some Tassellation Stuff}
+\label{sec:tass_stuff}
 
@@
 
-\begin{figure}[ht]%[htbp]
-\begin{center}
-\begin{center}\includegraphics[width = 1\textwidth]{texas_model.jpg}\end{center}
-\caption{TNA dual graph of the MLK memorial blah blah\ldots}
-\label{fig:texas_model}
-\end{center}
-\end{figure}
+\begin{equation}
+X = x_{1},x_{2},x_{3},x_{4},x_{5},x_{6},x_{7},x_{8}
+\end{equation}
 
@@
 
- This is some pseudocode example from an awesome PhD thesis by Tom\'{a}s
- M\'{e}ndez Echenagucia:\\
-
- \begin{algorithmic}[1]
-\State Generate a random and coded population 
-\For{$i\gets 1$, number of Generations}
-    \State Decode and Scale the population
-    \State Calculate fitness values for the population
-    \State Run Selection operation
-    \State Run Crossover operation
-    \State Run Mutation operation
-    \State Perform Exit condition test
-
-    \If{Exit condition test = \texttt{True}}
-        \State \texttt{Exit} GA
-    \ElsIf{Exit condition test = \texttt{False}}
-        \State Continue to the next Generation
-    \EndIf
-\EndFor
-\end{algorithmic}
-
-I told you it was awesome.
+  Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
+ incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
